### 2018-02-28 paper/compute.tex
Add first draft of the paper

\section{Compute}\label{sec-compute}

\begin{figure}
\begin{minted}{haskell}
type IdentityCompute    k v = forall f.                  (k -> f v) -> k -> f v
type FunctorialCompute  k v = forall f. Functor     f => (k -> f v) -> k -> f v
type ApplicativeCompute k v = forall f. Applicative f => (k -> f v) -> k -> f v
type AlternativeCompute k v = forall f. Alternative f => (k -> f v) -> k -> f v
type MonadicCompute     k v = forall m. Monad       m => (k -> m v) -> k -> m v
type MonadPlusedCompute k v = forall m. MonadPlus   m => (k -> m v) -> k -> m v
\end{minted}
\caption{Compute abstractions}\label{fig-compute}
\end{figure}

Consider abstractions in Fig.~\ref{fig-compute}.