### 2018-03-07 paper/engineering.tex

\clearpage
\section{Engineering aspects}\label{sec-engineering}

\begin{itemize}
    \item Sandboxing
    \item Non-determinism, specifically, it's useless form. This doesn't
          include \Excel's \textsf{RANDBETWEEN} function, which is useful.
          To define correctness of non-deterministic build systems, one needs to
          switch from \hs{Applicative} and \hs{Monad} abstractions to
          \hs{Alternative} and \hs{MonadPlus}, respectively, but this is not
          required for actual implementations, most of which can happily accept
          non-deterministic results (with a notable exception of \Buck).
    \item Concurrency
    \item Cloud
\end{itemize}