### 2018-03-07 paper/related.tex

\clearpage
\section{Related work}\label{sec-related}

\begin{itemize}
    \item Memoization
    \item Self-adjusting computation
    \item Lens-like formulation
\end{itemize}