### 2018-03-08 paper/engineering.tex
Minor revision

\clearpage
\section{Engineering aspects}\label{sec-engineering}

Or ``Build systems in the real world'', or ``Crazy things section''...

\begin{itemize}
    \item Sandboxing
    \item Non-determinism, specifically, its useless form. This doesn't
          include \Excel's \textsf{RANDBETWEEN} function, which is useful.
          To define correctness of non-deterministic build systems, one needs to
          switch from \hs{Applicative} and \hs{Monad} abstractions to
          \hs{Alternative} and \hs{MonadPlus}, respectively, but this is not
          required for actual implementations, most of which can happily accept
          non-deterministic results (with a notable exception of \Buck).
    \item Concurrency: nothing fundamentally difficult and fits our theory, but
          hard to get right in practice.
    \item Cloud aspects: eviction/garbage collection, `frankenbuilds', etc.
    \item Iterative computations, e.g. LaTeX.
    \item Polymorphic keys and values: tasks with multiple outputs, e.g. GHC's
          \cmd{hi} and \cmd{o} files, phony tasks, etc.
\end{itemize}