### 2018-03-09 paper/related.tex
Finish Shake, Bazel, add overview draft

\clearpage
\section{Related work}\label{sec-related}

\begin{itemize}
    \item Memoization
    \item Self-adjusting computation
    \item Lens-like formulation
    \item Other build systems: \Ninja, \cmd{nix},
          cloud (Blaze -- first? \Buck, Pants, CloudBuild, XXX),
          Shake derivatives (Jane Street's \cmd{Jenga}, Apple's \cmd{swift-llbuild}).
\end{itemize}

John Ericson suggested in a blog comment that nix may be somewhat monadic, see:
\url{https://blogs.ncl.ac.uk/andreymokhov/cloud-and-dynamic-builds/\#comment-1849}.