### 2018-03-09 paper/build.tex
Restructure the paper

\section{Build systems, abstractly}\label{sec-build}

In this section we further analyse the \hs{Build i c k v} abstraction, in
particular, we analyse two main aspects of every build system: the metadata
\hs{i} it persistently stores between builds, and the algorithm it uses to
compute values in the dependency order.

\subsection{Metadata: dirty bit, dependency graph, cache}\label{sec-build-metadata}

...

\subsection{Algorithm: topological, reordering, recursive}\label{sec-build-algorithm}

...