### 2018-03-13 paper/3-compute.tex
Stuff from Simon

\section{Build systems, abstractly}\label{sec-compute}

This section presents purely functional abstractions that allow us to express
all the intricacies of build systems discussed in the previous
section~\S\ref{sec-background} and design complex build systems from simple
primitives. Specifically, we present the \emph{compute} and \emph{build}
abstractions, which are described in
Sections~\ref{sec-task} and \ref{sec-general-build} respectively.

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
type Task c k v = @\std{forall}@ f. c f => (k -> f v) -> k -> Maybe (f v)
type Build c i k v = Task c k v -> k -> Store i k v -> Store i k v
\end{minted}
\vspace{1mm}

\hs{Build} corresponds to the algorithm used for bringing a key-value store
\store to a consistent state by recomputing out-of-date values in an
appropriate order. We explain the build type and use it to develop a formal
definition of build systems in~\S\ref{sec-general-build}.
Sections~\S\ref{sec-build} and~\S\ref{sec-implementations} scrutinise the build
abstraction further and provide concrete implementations for several build
systems.

\subsection{Common vocabulary for build systems}
\label{sec-vocabulary}

\begin{figure}
\begin{minted}{haskell}
-- Stores
data Store i k v   -- Abstract
getInfo  :: Store i k v -> i
putInfo  :: Store i k v -> i -> Store i k v
getValue :: Store i k v -> k -> v
putValue :: Hashable v => Store i k v -> k -> v -> Store i k v
getHash  :: Hashable v => Store i k v -> k -> Hash

-- Hashing
hash :: Hashable a => a -> Hash

-- Applicative
(<*>) :: Applicative f => f (a->b) -> f a -> f b  -- Left-associative
(<$>) :: Applicative f => (a->b) -> f a -> f b    -- Left-associative
pure  :: Applicative f => a -> f a

-- State monad
data State s
instance Applicative State
instance Monad       State
evalState :: State s a -> s -> s
getState  :: State s s
\end{minted}
\caption{Signatures of library functions}
\label{fig-lib-funs}
\end{figure}

We start by establishing a common vocabulary for build systems that allows
us to abstract away from specific application domains, such as software
development or spreadsheets.
% We will use this vocabulary throughout the rest of the paper.

A \emph{build system} takes a \emph{task description}, a target \emph{key},
and a \emph{store}, and returns a new store in which the target key is
up to date.  The store is a key/value mapping, togther with
\emph{persistent build state} that is maintained by the build system itself.

In software build systems, keys are typically filenames,
e.g.~\cmd{main.c}, and values are file contents (a C program source
code in this case). In spreadsheets, keys are cell names,
e.g. \cmd{A1}, and values are numbers, text, etc. that are typically
displayed inside cells. In our Haskell code, we will use type
variables \hs{k} and \hs{v} to denote keys and values, respectively.

Some values must be provided by the user as \emph{input}. For example,
\cmd{main.c} can be edited by the user who relies on the build system to
compile it into \cmd{main.o} and subsequently \cmd{main.exe}. End build products,
such as \cmd{main.exe}, are \emph{output} values. All other values (in this case
\cmd{main.o}) are \emph{intermediate}; they are not interesting for the user
but are produced in the process of turning inputs into outputs.

The types of the library functions we use in this paper is
given in Figure~\ref{fig-lib-funs}.
We use a cryptographic \emph{hash function} \hs{hash :: Hashable v =>
v -> Hash} for efficient tracking and sharing of~build results, where
\hs{Hash} is an abstract data type equipped with the equality test.

A \emph{store}, of type \hs{Store i k v}, associates keys to values.
It is convenient to assume that a
store is total function \store, i.e. it contains a value for every possible key.
We therefore also assume that the type of values is capable of encoding values
corresponding to non-existent keys (missing files or empty cells). This, in
particular, suggests that it is possible to \emph{depend on the absence of a
value}, which is a useful feature for build systems. In addition to usual
read and write operations, some stores also support a dedicated \emph{read hash}
operation, which is implemented more efficiently than by reading and hashing the
result. For example, GVFS (Git Virtual File System)~\cite{gvfs} stores file
hashes and downloads file contents only on demand -- an important feature for
cloud build systems.

The store also contains a single persistent value, of type \hs{i}, representing
the persistent state maintained by the build system itself -- its ``memory''.

\subsection{The Task abstraction}\label{sec-task}

Our first key idea is the task-description abstraction:
\begin{minted}[xleftmargin=10pt]{haskell}
type Task c k v = @\std{forall}@ f. c f => (k -> f v) -> k -> Maybe (f v)
\end{minted}
This highly-abstracted type is best introduced through an example.
Consider this Excel spreadsheet (c.f. \S\ref{sec-background-excel}):
\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{bash}
  A1: 10     B1: =A1 + A2
  A2: 20     B2: =B1 * 2
\end{minted}
\vspace{1mm}
Here cell \hs{A1} contains the value \hs{10} while cell \hs{B1} contains
the formula \hs{=A1+A2}.
We can represent the formulae of of this spreadsheet
with the following task description
\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
sprsh1 :: Compute Applicative String Integer
sprsh1 get "B1" = Just ((+) <$> get "A1" <*> "A2")
sprsh1 get "B2" = Just ((* 2) <$> get "B1")
sprsh1 _   _    = Nothing
\end{minted}
\vspace{1mm}
Here we instantiate keys \hs{k} with \hs{String}, and values \hs{v} with \hs{Integer}.
(Real spreadsheet cells would contain a wider range of values, of course.)
The task description \hs{sprsh1} embodies all the \emph{formulae} of the spreadsheet,
but not the constant values.  Like every \hs{Task}, \hs{sprsh1} is given
a \hs{get} function and a key. It pattern-matches on the key to see if it has
a task description (in the Excel case, a formula)
for it.  If not, it returns \hs{Nothing}, indicating that there is no formula in
that cell.  If there is a formula in the cell, it computes
the value of the formula, using \hs{get} to find the value of any keys on
which it depends.

The code to ``compute the value of a formula'' in \hs{sprsh1} looks a bit mysterious
because it takes place in an \hs{Applicative} computation \cite{mcbride2008applicative}
-- the relevant type signatures
are given in Figure~\ref{fig-lib-funs}.  We will return to the need for this
generality in
Section~\ref{sec-deps}\footnote{Readers familiar with \emph{lenses} or \emph{profunctor optics}
might recognise a familiar pattern. We discuss this in~\S\ref{sec-related-optics}.}.

For now, we content ourselves with observing that a task description,
of type \hs{Task c k v}, is completely isolated from the world of
compilers, calculation chains, file systems, caches, and all other
complexities of real build systems.  It just computes a single output, in
a side-effect-free way, using a call-back (\hs{get}) to get the values
of its dependencies.

\subsection{The Build abstraction}\label{sec-general-build}

Next comes our second key abstraction: a build system:
\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
type Build c i k v = Task c k v -> k -> Store i k v -> Store i k v
\end{minted}
\vspace{1mm}
The signature is very straightforward.  Given a task description, a target key,
and a store, the build system returns a new store in which the value of the
target key is up-to-date.  (What exactly does ``up to date'' mean?  We answer
that precisely in \S\ref{sec-build-correctness}.

\emph{SLPJ: is this right?}
\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
dumbBuild :: Build Monad () k v
dumbBuild task k store = execState (go k) store
  where
    go :: k -> State s v
    go k = case (task go k) of
             Nothing  -> do { s <- getState; return (getValue s k) }
             Just act -> do { v <- act; updState (\s -> putStore s k v); return v }
\end{minted}
\vspace{1mm}
\emph{SLPJ: explain}

\subsection{Correctness of a build system} \label{sec-build-correctness}

We can now say what it means for a build system to be \emph{correct}, something
that is seldom stated formally:

\definition[Correctness]{Suppose \hs{build :: Build c i k v} is a build system,
\hs{task :: Task c k v} describes an acyclic collection of build tasks, \hs{key} is an output
key, \hs{store} be a store.
Then the \hs{result}ing store \hs{result = build compute key store}
is~\emph{correct} if there exists a store \hs{ideal}, such that

\begin{itemize}
    \item The \hs{ideal} store is \emph{consistent} with \hs{task}, i.e.
    for all possible non-input keys \hs{k}, the result of executing the
    \hs{compute} matches the value in the \hs{ideal} store:
    \[
    \hs{execute}~\hs{ideal}~\hs{k}~\hs{==}~\hs{Just}~\hs{(}\hs{ideal}~\hs{k)}.
    \]
    \item The \hs{result} and \hs{ideal} agree on the output \hs{key}, i.e.
    \hs{result}~\hs{key}~\hs{==}~\hs{ideal}~\hs{key}. In other words, we require
    that the output \hs{key} has the correct value, but do not impose any
    restrictions on intermediate keys, therefore permitting shallow cloud builds.
    \item The \hs{store}, \hs{ideal} and \hs{result} all agree on the inputs
    that belong to the transitive closure of \hs{key}'s dependencies. This means
    that (i) no inputs were corrupted during the build, and (ii)~the \hs{ideal}
    store has constraints not only on the output, but also on the inputs.
\end{itemize}
}
\label{def-correct}

\subsection{Computing dependencies}

The \hs{Compute} type and the example above might seem mysterious. Why insist on
\hs{Compute} to be polymorphic in \hs{f}? Why choose the \hs{Applicative}
constraint?

The answer to both questions is that we can statically analyse an applicative
compute and extract all dependencies of a key using the following function:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
import Data.Functor.Const
\end{minted}
\vspace{0.5mm}
\begin{minted}[xleftmargin=10pt]{haskell}
dependencies :: Compute Applicative k v -> k -> [k]
dependencies compute = maybe [@@] getConst . compute (\k -> Const [k])
\end{minted}
\vspace{1mm}

\noindent
We can use this function to extract dependencies of \hs{add}, as shown in the
following GHCi session:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
@\ghci@ dependencies add "A1"
[@@]
\end{minted}
\begin{minted}[xleftmargin=10pt]{haskell}
@\ghci@ dependencies add "B1"
["A1", "A2"]
\end{minted}
\vspace{1mm}

\noindent
It is important to note that we extract the dependencies without providing any
values to the compute. This guarantees that no actual computation is performed
during the analysis -- we cannot execute the function \hs{(+)} if we have no
values for it! Static analysis of applicative computations is not new and has
been used to similar effect before, e.g. see~\cite{free-applicatives}.

To understand how \hs{dependencies} works, let us examine its type more closely,
expanding the type synonym \hs{Compute}:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
(@\std{forall}@ f. Applicative f => (k -> f v) -> k -> Maybe (f v)) -> k -> [k]
\end{minted}
\vspace{1mm}

\noindent
Since the first argument is polymorphic in \hs{f}, we can choose any \hs{f} that
satisfies the \hs{Applicative} constraint. A good choice in this case is the
applicative functor \hs{Const}:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
newtype Const m a = Const { getConst :: m }
\end{minted}
\vspace{0.5mm}
\begin{minted}[xleftmargin=10pt]{haskell}
instance Functor (Const m) where
    fmap _ (Const m) = Const m
\end{minted}
\vspace{0.5mm}
\begin{minted}[xleftmargin=10pt]{haskell}
instance Monoid m => Applicative (Const m) where
    pure _              = Const mempty
    Const x <*> Const y = Const (x <> y)
\end{minted}
\vspace{1mm}

\noindent
A value of type \hs{Const}~\hs{[}\hs{k]}~\hs{v} contains no values \hs{v} but
allows us to record all invocations of the custom \hs{fetch} function that we
feed to the given compute: \hs{fetch k = Const [k]}. The compute returns a
value of type \hs{Maybe}~\hs{(Const}~\hs{[}\hs{k]}~\hs{v)} that can be
unwrapped using \hs{maybe [@@] getConst}, i.e. returning the empty list of
dependencies in case of an input.

We can also \hs{execute} a compute on a given store. This relies on a similar
transformation, now based on the \hs{Identity} monad (see
Appendix~\S\ref{sec-appendix-transformers} for the implementation of
\hs{Identity}):

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
import Data.Functor.Identity
\end{minted}
\vspace{0.5mm}
\begin{minted}[xleftmargin=10pt]{haskell}
execute :: Compute Monad k v -> (k -> v) -> k -> Maybe v
execute compute store = fmap runIdentity . compute (pure . store)
\end{minted}
\vspace{1mm}

\noindent
Note that \hs{execute} can accept both an applicative as well as a \emph{monadic
compute} introduced in~\S\ref{sec-compute-monad}. Below we run \hs{execute} on
the above \hs{add} example. Note that \hs{execute} returns \hs{Nothing} for
inputs.

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
@\ghci@ store key = if key == "A1" then 10 else 20
@\ghci@ execute add store "A1"
Nothing
\end{minted}
\begin{minted}[xleftmargin=10pt]{haskell}
@\ghci@ execute add store "B1"
Just 30
\end{minted}

\subsection{Monadic compute}\label{sec-compute-monad}

As discussed in~\S\ref{sec-background}, some build tasks have dynamic
dependencies, which are determined by values of intermediate computations. In
our framework, such build tasks correspond to \emph{monadic compute}, i.e.
the type \hs{Compute}~\hs{Monad}~\hs{k}~\hs{v}. Recall the following spreadsheet
example:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{bash}
A1 = 10
A2 = 20
B1 = IF(C1=1,A1,A2)
C1 = 1
\end{minted}
\vspace{1mm}

\noindent
The task corresponding to the cell \cmd{B1} statically depends on \cmd{C1} and
dynamically depends on either \cmd{A1} or \cmd{A2}. We can express this using
our compute abstraction as follows\footnote{The spreadsheet example that uses
the \hs{INDIRECT} function can be expressed very similarly: simply replace the
last line of the \hs{select} function with \hs{fetch ("A" ++ show c1)}.}:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
select :: Compute Monad String Integer
select fetch key | key /= "B1" = Nothing
                 | otherwise   = Just $ do
                     c1 <- fetch "C1"
                     if c1 == 1 then fetch "A1" else fetch "A2"
\end{minted}
\vspace{1mm}

We cannot statically analyse \hs{select} using the function \hs{dependencies}
from the previous section, but we can still \hs{execute} it on a given store:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
@\ghci@ store key = if key == "A1" then 10 else 20
@\ghci@ execute select store "B1"
Just 20
\end{minted}
\vspace{1mm}

\noindent
In the example above $\cmd{C1} \neq 1$, hence the \hs{else} branch is taken.
To extract the dependencies of \hs{select} we introduce the function \hs{track}:
a combination of \hs{execute} and \hs{dependencies} that takes the store~\store
as input and computes both the resulting value and the list of its dependencies:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
import Control.Monad.Writer
\end{minted}
\vspace{0.5mm}
\begin{minted}[xleftmargin=10pt]{haskell}
track :: Compute Monad k v -> (k -> v) -> k -> Maybe (v, [k])
track compute store = fmap runWriter . compute (\k -> writer (store k, [k]))
\end{minted}
\vspace{1mm}

\noindent
The implementation uses the \hs{Writer} monad (\S\ref{sec-appendix-transformers}):
we feed \hs{fetch k = writer (store k, [k])} to the compute, which in addition
to fetching a value from the store, tracks the corresponding key. We can
\hs{track} both applicative and monadic compute:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
@\ghci@ store key = if key == "A1" then 10 else 20
@\ghci@ track add store "B1"
Just (30,["A1","A2"])
\end{minted}
\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
@\ghci@ track select store "B1"
Just (20,["C1","A2"])
\end{minted}
\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
@\ghci@ store' key = if key == "A1" then 10 else 1
@\ghci@ track select store' "B1"
Just (10,["C1","A1"])
\end{minted}
\vspace{1mm}

\noindent
As expected, the dependencies of \hs{select} are determined by the contents of
the store.

% \subsection{Recognising an applicative compute wearing a monadic hat}

% Although dynamic dependencies are very useful, they typically represent only a
% small fraction of all dependencies. One can exploit this in a build system by
% recognising applicative computations even though they have the type
% \hs{Compute}~\hs{Monad} and treating them specially.

% Consider the following data type.

% \subsection{Free build structures}\label{sec-free-build}

% \begin{minted}{haskell}
% data Depend k v r = Depend [k] ([v] -> r)

% data Depends k v r = Depends [k] ([v] -> Depends k v r)
%                    | Done r
% \end{minted}