### 2018-03-15 paper/8-conclusions.tex

\section{Conclusions}\label{sec-conclusions}

We have investigated multiple build systems, showing how their properties are consequences of two implementation choices: what order you build in and whether you decide to rebuild. By first decomposing the pieces, we show how to recompose the pieces to find new points in the design space. In particular, a simple recombination leads to a design for a monadic cloud build system. Armed with that blueprint we hope to actually implement such a system as future work.