### 2018-05-02 paper/response.md
Draft a response

# Reviewer A

Thank you!

# Reviewer B

We disagree about correctness vs minimality: our main focus is correctness.

# Reviewer C

> the code of each build system is quite different from each other

The framework does provide reusable "schedulers" (topological, recursive, etc.)
And in the final version we will also provide reusable "rebuilders".

> Build abstraction is trivial

The whole point of the paper is to ignore many crucially important and
substantial engineering aspects of build systems and bring out more subtle
design choices.

Parallelism is an example.