### 2018-06-30 paper/5-implementations.tex
Simplify build system implementations

\section{Build systems, concretely}\label{sec-implementations}

In the previous sections we discussed the types of build systems, and how they
can be broken down into two main components: a scheduler and a rebuilder.
In this section we make this abstract distinction concrete, by
implementing a number of build systems as a composition of a scheduler
and a rebuilder.  Our design is expressed by these types:
\begin{minted}[xleftmargin=10pt]{haskell}
type Scheduler c i ir k v = Rebuilder c ir k v -> Build c i k v
type Rebuilder c   ir k v = k -> v -> Task c k v -> Task (MonadState ir) k v
\end{minted}
So a \hs{Scheduler} is just a function that takes a \hs{Rebuilder} and uses
it to construct \hs{Build} system, by choosing which keys to rebuild in which order.
The \hs{Rebuilder} makes uses of the persistent build information \hs{ir}, while
the scheduler might augment that with further persistent information of its own,
yielding \hs{i}.

A \hs{Rebuilder} is a function which the \hs{Scheduler} uses to rebuild a
particular key, passing to it the key, the key's current value, and a \hs{Task}
that can compute the value of the key if necessary. It uses the persistent build
information \hs{ir} (carried by the state monad) to decide whether to rebuild
the value. If doing so is unnecessary, it returns the current value; otherwise
it runs the supplied \hs{Task} to rebuild it. In both cases it can choose to
update the persistent build information \hs{ir} to reflect what happened. So a
\hs{Rebuilder} wraps a \hs{Task}~\hs{c}~\hs{k}~\hs{v}, which unconditionally
rebuilds the key, to make a \hs{Task}~\hs{(MonadState}~\hs{ir)}~\hs{k}~\hs{v},
which rebuilds the key only if necessary, and does the necessary book-keeping.
Note that the resulting \hs{Task} is always monadic; static dependency analysis
can be performed on the original \hs{Task}~\hs{Applicative} if need be.

These two abstractions are the key to modularity: \emph{we can combine any
scheduler with any rebuilder, and obtain a correct build system}.
In this section we will write a scheduler for each column of
Table~\ref{tab-build-systems}, and a rebuilder for each row; then combine them
to obtain the build systems in the table's body.

\subsection{\Make}\label{sec-implementation-make}

An implementation of \Make using our framework is shown in
Fig.~\ref{fig-make-implementation}. As promised, its definition
is just the application of a \hs{Scheduler}, \hs{topological},
to a \hs{Rebuilder}, \hs{modTimeRebuilder}.
We discuss each component in turn, starting with the rebuilder.

The \hs{modTimeRebuilder} uses the pair
\hs{MakeInfo}~\hs{k}~\hs{=}~\hs{(@@now,}~\hs{modTimes)} as persistent
build information, carried by a \hs{MonadState} monad. This
\hs{MakeInfo} comprises the \emph{current time} \hs{now} and the map
\hs{modTimes}~\hs{::}~\hs{Map}~\hs{k}~\hs{Time} of \emph{file
modification times}. The rebuilder receives three arguments: a
\hs{key}, its current \hs{value}, and the \hs{task} of type
\hs{Task}~\hs{Applicative}~\hs{k}~\hs{v} that can be used to rebuild
the \hs{key} if necessary.  The rebuilder first decides if the
\hs{key} is \hs{dirty} by consulting \hs{modTimes}: if the key is not
found, that must mean it has never been built before; otherwise
\hs{modTimeRebuilder} can see if any of the \hs{task}'s dependencies
(computed by \hs{dependencies}) are out of date. If the key is
\hs{dirty}, we use \hs{run}~\hs{task} to rebuild it, and update the state with
the new modification time of the key\footnote{The real \Make
relies on the file system to track modification times, but we prefer
to make it explicit here.}; otherwise we can just return the current value.
We assume that the external system, which invokes the build system, updates
\hs{MakeInfo}~\hs{k} reflecting any file changes between successive builds.

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Make build system; stores current time and file modification times
type Time       = Integer
type MakeInfo k = (Time, Map k Time)
\end{minted}
\vspace{0mm}
\begin{minted}[fontsize=\small]{haskell}
make :: Ord k => Build Applicative (MakeInfo k) k v
make = topological modTimeRebuilder
\end{minted}
\vspace{0mm}
\begin{minted}[fontsize=\small]{haskell}
-- A task rebuilder based on file modification times
modTimeRebuilder :: Ord k => Rebuilder Applicative (MakeInfo k) k v
modTimeRebuilder key value task = Task $ \fetch -> do
    (now, modTimes) <- get
    let dirty = case Map.lookup key modTimes of
            Nothing -> True
            time -> any (\d -> Map.lookup d modTimes > time) (dependencies task)
    if not dirty then return value else do
        put (now + 1, Map.insert key now modTimes)
        run task fetch
\end{minted}
\vspace{0mm}
\begin{minted}[fontsize=\small]{haskell}
-- A topological task scheduler
topological :: Ord k => Scheduler Applicative i i k v
topological rebuilder tasks target = execState $ forM_ order $ \key -> case tasks key of
    Nothing   -> return ()
    Just task -> do
        store <- get
        let newTask :: Task (MonadState i) k v
            newTask = rebuilder key (getValue key store) task
            fetch :: k -> State i v
            fetch k = return (getValue k store)
        newValue <- liftStore (run newTask fetch)
        modify $ putValue key newValue
  where
    deps k = case tasks k of { Nothing -> [@@]; Just task -> dependencies task }
    order  = fromJust $ topSort (reachable deps target) -- Assumes acyclic dependencies
\end{minted}
\vspace{0mm}
\begin{minted}[fontsize=\small]{haskell}
-- Standard graph algorithms (implementation omitted)
reachable :: Ord k => (k -> [k]) -> k -> Graph k
topSort   :: Ord k => Graph k -> Maybe [k] -- Returns Nothing on a cyclic graph
\end{minted}
\vspace{0mm}
\begin{minted}[fontsize=\small]{haskell}
-- Expand the scope of visibility of a stateful computation (implementation omitted)
liftStore :: State i a -> State (Store i k v) a
\end{minted}
\vspace{-3mm}
\caption{An implementation of \Make using our framework.}\label{fig-make-implementation}
\vspace{-3mm}
\end{figure}

\Make's scheduler, \hs{topological}, processes keys in a linear \hs{order} based on a
topological sort of the statically known dependency graph
(see~\S\ref{sec-parallelism} for parallel \Make). Our definition in
Fig.~\ref{fig-make-implementation} is polymorphic with respect to the type of
build information \hs{i} and is therefore compatible with any applicative
\hs{rebuilder}. The scheduler calls the supplied \hs{rebuilder} on every
\hs{key} in the \hs{order}, and runs the obtained \hs{newTask} to compute the
\hs{newValue} and updated build information \hs{newInfo}, which are subsequently
written to the store. The computation runs in the
\hs{State}~\hs{(@@Store}~\hs{i}~\hs{k}~\hs{v)} monad, similar to the \hs{busy}
build system (\S\ref{sec-general-build}), however since \hs{newTask} has access
only to the \hs{i} part of the \hs{Store}~\hs{i}~\hs{k}~\hs{v}, we need to use
\hs{runState} to execute it. The \hs{newTask} finds values of the \hs{key}'s
dependencies via the \hs{fetch} callback, which is defined to directly read the
\hs{store}.

The \hs{where} clause corresponds to the pre-processing stage, which uses the
function \hs{dependencies}, defined in~\S\ref{sec-deps}, to extract static
dependencies from the provided applicative \hs{task}. We compute the linear
processing \hs{order} by constructing the graph of keys \hs{reachable} from the
\hs{target} via dependencies, and performing the topological sort of the result.
We omit implementation of textbook graph algorithms
\hs{reachable} and \hs{topSort}, e.g. see~\cite{cormen2001introduction}.

Note that the function \hs{dependencies} can only be applied to applicative
tasks, which restricts \Make to static dependencies, as reflected in the
type~\hs{Build}~\hs{Applicative}. Moreover, any other build system that uses
the \hs{topological} scheduler will also inherit the same restriction.

\subsection{\Excel}\label{sec-implementation-excel}

Our \Excel's model is a composition of the \hs{restarting} scheduler and the
\hs{approximateRebuilder}, as shown in Fig.~\ref{fig-excel-implementation}.
The persistent build information \hs{ExcelInfo}~\hs{k} is a triple:

\begin{itemize}
    \item A set of dirty keys \hs{Set}~\hs{k}, i.e. all cells whose contents
          has changed since the previous build, including input cells and
          formulas.
    \item A map \hs{Map}~\hs{k}~\hs{[@@k]} of known upper bounds of
          dependencies. If there is an entry for a key, it is a conservative
          approximation of its dependencies (e.g. the dependencies of the
          formula \cmd{IF(C1=1,B2,A2)} are approximated by the list
          \hs{[}\cmd{C1}\hs{,}\cmd{B2}\hs{,}\cmd{A2}\hs{]}). Otherwise, there is
          no upper bound (e.g. because the key uses an \cmd{INDIRECT} reference)
          and the key is assumed to be dirty.
    \item A calculation \hs{Chain} recorded in the previous build
          (\S\ref{sec-background-excel}), represented by a list of keys \hs{[@@k]}.
\end{itemize}

The \hs{approximateRebuilder} uses the first two elements of the above triple
to decide whether a given \hs{key} is \hs{dirty} and hence needs to be rebuilt.
The logic is similar to that of \Make's \hs{modTimeRebuilder}, but instead of
relying on file modification times, \Excel stores the information about dirty
keys more directly. Furthermore, \Excel's dependency approximation makes it
possible to build tasks with dynamic dependencies: a \hs{key} is declared
\hs{dirty} if any of its approximate dependencies is in the \hs{dirtyKeys} set,
or if there is no known upper bound on its dependencies, i.e. if the lookup in
the map \hs{deps}~\hs{::}~\hs{Map}~\hs{k}~\hs{[@@k]} fails. Note that when a
\hs{key} is rebuilt it is added to the set \hs{dirtyKeys} to trigger further
rebuilds.

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Excel build system; stores a dirty bit per key and calc chain
type Chain k = [k]
type ExcelInfo k = (k -> Bool, Chain k)
\end{minted}
\vspace{0mm}
\begin{minted}[fontsize=\small]{haskell}
excel :: Ord k => Build Monad (ExcelInfo k) k v
excel = restarting dirtyBitRebuilder
\end{minted}
\vspace{0mm}
\begin{minted}[fontsize=\small]{haskell}
-- A task rebuilder based on dirty bits
dirtyBitRebuilder :: Rebuilder Monad (k -> Bool) k v
dirtyBitRebuilder key value task = Task $ \fetch -> do
    isDirty <- get
    if isDirty key then run task fetch else return value
\end{minted}
\vspace{0mm}
\begin{minted}[fontsize=\small]{haskell}
-- A restarting task scheduler
restarting :: Ord k => Scheduler Monad (ir, Chain k) ir k v
restarting rebuilder tasks target = execState $ do
    chain    <- gets (snd . getInfo)
    newChain <- liftChain $ go Set.empty $ chain ++ [target | target `notElem` chain]
    modify . mapInfo $ \(ir, _) -> (ir, newChain)
  where
    go :: Set k -> Chain k -> State (Store ir k v) (Chain k)
    go _    []       = return []
    go done (key:ks) = case tasks key of
        Nothing -> (key :) <$> go (Set.insert key done) ks
        Just task -> do
            store <- get
            let value = getValue key store
                newTask :: Task (MonadState ir) k (Either k v)
                newTask = try $ rebuilder key value task
                fetch :: k -> State ir (Either k v)
                fetch k | k `Set.member` done = return $ Right (getValue k store)
                        | otherwise           = return $ Left k
            result <- liftStore (run newTask fetch)
            case result of
                Left dep -> go done $ dep : filter (/= dep) ks ++ [key]
                Right newValue -> do
                    modify $ updateValue key value newValue
                    (key :) <$> go (Set.insert key done) ks
\end{minted}
\vspace{0mm}
\begin{minted}[fontsize=\small]{haskell}
-- Convert a total task into a task that accepts a partial fetch callback
try :: Task (MonadState i) k v -> Task (MonadState i) k (Either e v)
try task = Task $ \fetch -> runExceptT $ run task (ExceptT . fetch)
\end{minted}
\vspace{0mm}
\begin{minted}[fontsize=\small]{haskell}
-- Expand the scope of visibility of a stateful computation (implementation omitted)
liftChain :: State (Store ir k v) a -> State (Store (ir, Chain [k]) k v) a
\end{minted}
\vspace{-3mm}
\caption{An implementation of \Excel using our framework.}\label{fig-excel-implementation}
\vspace{-5mm}
\end{figure}

\Excel's \hs{restarting} scheduler processes keys in the order specified by the
calculation \hs{chain}. During the build, it constructs a \hs{newChain} for the
next build and maintains a set of keys \hs{done} that have been processed. For
each non-input \hs{key}, the scheduler tries to rebuild it using a partial
\hs{fetch} callback that returns \hs{Either}~\hs{k}~\hs{v} instead of \hs{v}.
The callback is defined to fail with \hs{Left}~\hs{dep} when asked for the value
of a dependency \hs{dep} that has not yet been processed (and hence may
potentially be dirty); otherwise it returns the current value of the dependency
by looking it up in the \hs{store}.

After the \hs{newTask} is executed with \hs{runState} there are two cases to
consider:

\begin{itemize}
    \item The \hs{newTask} has failed, because one of its dependencies \hs{dep}
    has not yet been processed. This indicates that the calculation \hs{chain}
    from the previous build is incorrect and needs to be adjusted by making sure
    it contains \hs{dep} and moving the \hs{key} to the end, so that its build
    can be restarted after the \hs{dep} is ready.
    \item The \hs{newTask} succeeded. The resulting \hs{newValue} and
    \hs{newInfo} are written to the store, the \hs{key} is marked as \hs{done},
    and \Excel continues to build the \hs{chain}.
\end{itemize}

Note that the task returned by the \hs{rebuilder} expects a total callback
function and cannot be directly executed with the partial callback \hs{fetch}.
We fix the mismatch with the function \hs{try} that relies on the standard
monad transformer \hs{ExceptT} from the \cmd{transformers} library.

\vspace{-1mm}
\subsection{\Shake}\label{sec-implementation-shake}
\vspace{-1mm}

Our model of \Shake is shown in Fig.~\ref{fig-shake-implementation}. It stores
verifying traces \hs{VT}~\hs{k}~\hs{v} defined in~\S\ref{sec-verifying-traces}
as persistent build information and is composed of the \hs{suspending} scheduler
and the \hs{vtRebuilder}.

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Shake build system; stores verifying traces
shake :: (Ord k, Hashable v) => Build Monad (VT k v) k v
shake = suspending vtRebuilder
\end{minted}
\vspace{0mm}
\begin{minted}[fontsize=\small]{haskell}
-- A task rebuilder based on verifying traces
vtRebuilder :: (Eq k, Hashable v) => Rebuilder Monad (VT k v) k v
vtRebuilder key value task = Task $ \fetch -> do
    upToDate <- verifyVT key value (fmap hash . fetch) =<< get
    if upToDate then return value else do
        (newValue, deps) <- track task fetch
        put =<< recordVT key newValue deps (fmap hash . fetch) =<< get
        return newValue
\end{minted}
\vspace{0mm}
\begin{minted}[fontsize=\small]{haskell}
-- A suspending task scheduler
suspending :: Ord k => Scheduler Monad i i k v
suspending rebuilder tasks target store = fst $ execState (build target) (store, Set.empty)
  where
    build :: k -> State (Store i k v, Set k) (@@)
    build key = case tasks key of
        Nothing -> return (@@)
        Just task -> do
            done <- gets snd
            when (key `Set.notMember` done) $ do
                value <- gets (getValue key . fst)
                let newTask :: Task (MonadState i) k v
                    newTask = rebuilder key value task
                    fetch :: k -> StateT i (State (Store i k v, Set k)) v
                    fetch k = do lift (build k)                      -- build the key
                                 lift (gets (getInfo . fst)) >>= put -- save new traces
                                 lift (gets (getValue k . fst))      -- fetch the value
                newValue <- flattenInfo (run newTask fetch)
                modify $ \(s, d) -> (putValue key newValue s, Set.insert key d)
\end{minted}
\begin{minted}[fontsize=\small]{haskell}
-- Collapse two layers of stateful computations into one (implementation omitted)
flattenInfo :: StateT i (State (Store i k v, Set k)) a -> State (Store i k v, Set k) a
\end{minted}
\vspace{-3mm}
\caption{An implementation of \Shake using our framework.}\label{fig-shake-implementation}
\vspace{-6mm}
\end{figure}

The rebuilder executes the verification query \hs{verifyVT} to determine if the
\hs{key} is \hs{upToDate}. If it is, the rebuilder simply returns the \hs{key}'s
current \hs{value}. Otherwise it executes the \hs{task}, obtaining both a
\hs{newValue} and the \hs{key}'s dynamic dependencies \hs{deps} (see the
definition of \hs{track} in~\S\ref{sec-deps}), which are subsequently recorded
in a new verification trace using \hs{recordVT}.

The \hs{suspending} scheduler is based on a recursive function \hs{build} that
builds a given \hs{key}, making sure not to duplicate work when called on the
same \hs{key} again in future. To achieve that, it keeps track of keys that have
already been built in a set \hs{done}~\hs{::}~\hs{Set}~\hs{k}, running the
computation in the
\hs{State}~\hs{(@@Store}~\hs{i}~\hs{k}~\hs{v,}~\hs{Set}~\hs{k)} monad.

Given a non-input \hs{key} that has not yet been built, we use the supplied
\hs{rebuilder} to embed the build information~\hs{i} into the \hs{task}.
We then execute the obtained \hs{newTask} with the \hs{fetch} callback, which
performs the following three steps when given a dependency \hs{k}:

\begin{itemize}
    \item Build \hs{k} by recursively calling the \hs{build} function. Note
          that if \hs{k} is dirty, this will effectively suspend the build of
          the current \hs{key} until \hs{k} is brought up to date.
    \item Save any new traces that have been generated when building \hs{k}.
    \item Read the (up to date) value of \hs{k} from the store and return it.
\end{itemize}

These steps are run in the
\hs{StateT}~\hs{i}~\hs{(@@State}~\hs{(@@Store}~\hs{i}~\hs{k}~\hs{v,}~\hs{Set}~\hs{k))}
monad, because the \hs{newTask} has access only to the~\hs{i} part of the state;
we therefore need to use \hs{lift} when executing commands in the inner
\hs{State}~\hs{(@@Store}~\hs{i}~\hs{k}~\hs{v,}~\hs{Set}~\hs{k)} monad.
The pair \hs{(@@newValue,}~\hs{newInfo)} obtained by running the \hs{newTask} is
stored, and the \hs{key} is added to the set \hs{done}.

\subsection{Cloud build systems: \Bazel, \CloudBuild and \Cloud \Shake}\label{sec-implementation-cloud}

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Bazel build system; stores constructive traces
bazel :: (Ord k, Hashable v) => Build Monad (CT k v) k v
bazel = restarting2 ctRebuilder -- implementation of 'restarting2' is omitted (22 lines)
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- A rebuilder based on constructive traces
ctRebuilder :: (Eq k, Hashable v) => Rebuilder Monad (CT k v) k v
ctRebuilder key value task = Task $ \fetch -> do
    maybeCachedValue <- constructCT key value (fmap hash . fetch) =<< get
    case maybeCachedValue of
        Just cachedValue -> return cachedValue
        Nothing -> do
            (newValue, deps) <- track task fetch
            put =<< recordCT key newValue deps (fmap hash . fetch) =<< get
            return newValue
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- CloudBuild build system, implementation of 'topological' is given in Fig. 7
cloudBuild :: (Ord k, Hashable v) => Build Applicative (CT k v) k v
cloudBuild = topological (adaptRebuilder ctRebuilder)
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Convert a monadic rebuilder to the corresponding applicative one
adaptRebuilder :: Rebuilder Monad i k v -> Rebuilder Applicative i k v
adaptRebuilder rebuilder key value task = rebuilder key value $ Task $ run task
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Cloud Shake build system, implementation of 'suspending' is given in Fig. 9
cloudShake :: (Ord k, Hashable v) => Build Monad (CT k v) k v
cloudShake = suspending ctRebuilder
\end{minted}
\vspace{-2mm}
\caption{Implementations of \Bazel, \CloudBuild and \Cloud \Shake using our
framework; \hs{restarting2} differs from the restarting scheduler used by \Excel
(Fig.~\ref{fig-excel-implementation}) in that it does not rely on the
calculation chain.}
\label{fig-bazel-cloud-shake-implementation}
\vspace{-4mm}
\end{figure}

In this section we use constructive traces to implement cloud build systems.
Fig.~\ref{fig-bazel-cloud-shake-implementation} shows our models of \Bazel,
\CloudBuild and \Cloud \Shake, which both rely on the \hs{ctRebuilder} -- a
rebuilder based on constructive traces (\S\ref{sec-constructive-traces}).

The implementation of \hs{ctRebuilder} is analogous to that of \hs{vtRebuilder}
in Fig.~\ref{fig-shake-implementation}, but the \hs{verifyVT} query is replaced
with a more powerful query to \hs{constructCT} that can construct a resulting
\hs{cachedValue} by looking it up the cloud cache. If the cache lookup fails, we
fall back to executing the \hs{task}. The obtained \hs{newValue} and the
\hs{task}'s dependencies are recorded as a new constructive trace for future
use.

The \Bazel build system uses a restarting scheduler whose implementation we
omit. It is similar to \Excel's \hs{restarting} scheduler defined in
Fig.~\ref{fig-excel-implementation}, but instead of building keys in the order
specified by the persistently stored calculation chain, \Bazel uses a
\emph{build queue}. The build starts with the queue containing all dirty keys.
Similar to \Excel, the rebuilding of a key extracted from the queue may fail
because one of its dynamic dependencies is dirty. In this case the key is marked
as \emph{blocked} and its rebuilding is deferred. Whenever a key is successfully
rebuilt, all keys that were previously blocked on it are added back to the
queue, and their build is eventually restarted.

Note that although both our model and \Bazel's actual implementation supports
dynamic dependencies, it is currently not possible to define new monadic build
rules in the language available to users. Instead, users have to rely on a
(rich) collection of predefined built-in rules, which cover many but not all
practically important cases of dynamic dependencies.

By switching to the \hs{topological} scheduler, we obtain a model of
Microsoft's \CloudBuild~-- an applicative build system that combines
conventional scheduling of statically known directed acyclic graphs of
dependencies with constructive traces~\cite{esfahani2016cloudbuild}. Note that
we need to convert a monadic \hs{ctRebuilder} into an applicative one by
applying an adapter function \hs{adaptRebuilder}, which unwraps a given
\hs{Task Applicative} and wraps it into \hs{Task Monad}.

Using the abstractions built thus far, we have shown how to combine schedulers
with rebuilders to reproduce existing build systems. To us, the most interesting
build system as yet unavailable would combine a suspending scheduler with
constructive traces~--~providing a cloud-capable build system that is minimal,
and supports both early cutoff and monadic dependencies. Using our framework it
is possible to define and test such a system, which we call \Cloud \Shake. All
we need to do is replace \hs{restarting2} with \hs{suspending} in the definition
of \Bazel, as shown in Fig.~\ref{fig-bazel-cloud-shake-implementation}.

\textbf{TODO:} A real implementation of \Cloud \Shake is now available... It was
obtained by directly applying the presented ideas.

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Buck build system, implementation of 'topological' is given in Fig. 7
buck :: (Hashable k, Hashable v) => Build Applicative (DCT k v) k v
buck = topological (adaptRebuilder dctRebuilder)
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Nix build system, implementation of 'suspending' is given in Fig. 9
nix :: (Hashable k, Hashable v) => Build Monad (DCT k v) k v
nix = suspending dctRebuilder
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Rebuilder based on deterministic constructive traces
dctRebuilder :: (Hashable k, Hashable v) => Rebuilder Monad (DCT k v) k v
dctRebuilder key _value task = Task $ \fetch -> do
    dct <- get
    maybeCachedValue <- constructDCT key (fmap hash . fetch) dct
    case maybeCachedValue of
        Just cachedValue -> return cachedValue
        Nothing -> do
            (newValue, deps) <- track task fetch
            put =<< recordDCT key newValue deps (fmap hash . fetch) =<< get
            return newValue
\end{minted}
\vspace{-2mm}
\caption{Implementations of \Buck and \Nix using our framework.}
\label{fig-buck-nix-implementation}
\vspace{-2mm}
\end{figure}

\subsection{Buck and Nix}

Some build systems allow to exploit \emph{build task determinism} for efficient
fetching of build results from the cache: one can simply hash the values of all
transitive inputs of a key (which can be done efficiently by storing all hashes
in a Merkle tree) and download the matching value from the cache, without the
need to build any intermediate values. This can bring significant performance
benefits in practice, but sacrifices the early cutoff property: indeed, with
this approach addition of a new comment to one of the source files will
invalidate all dependent keys.

Fig.~\ref{fig-buck-nix-implementation} shows our models of \Buck and \Nix build
systems that use the \hs{dctRebuilder} -- a rebuilder based on deterministic
constructive traces (\S\ref{sec-deterministic-constructive-traces}). As one can
see, \hs{dctRebuilder} is almost identical to \hs{ctRebuilder}, but it ignores
the current \hs{value}, since the \hs{constructDCT} query only needs to examine
the transitive inputs of the given \hs{key}. \Buck uses the \hs{topological}
scheduler and is therefore an applicative build system, whereas \Nix uses the
\hs{suspending} scheduler and is monadic.