### 2019-04-28 papers/jfp/4-schedulers.tex

\section{Schedulers}\label{sec-scheduler}

The focus of this paper is on a variety of implementations of
\hs{Build}~\hs{c}~\hs{i}~\hs{k}~\hs{v}, given a \emph{user-supplied}
implementation of \hs{Tasks}~\hs{c}~\hs{k}~\hs{v}. That is, we are going to take
\hs{Tasks} as given from now on, and explore variants of \hs{Build}: first
abstractly (in this section and in~\S\ref{sec-rebuilder}) and then concretely
in~\S\ref{sec-implementations}.

As per the definition of minimality~\ref{def-minimal}, a minimal build
system must \textbf{rebuild only out-of-date keys} and at most once. The only
way to achieve the ``at most once'' requirement while producing a correct build
result (\S\ref{sec-build-correctness}) is to \textbf{build all keys in an
order that respects their dependencies}.

We have emboldened two different aspects above: the part of the build system
responsible for scheduling tasks in the dependency order (a ``scheduler'') can
be cleanly separated from the part responsible for deciding whether a key needs
to be rebuilt (a ``rebuilder''). In this section we discuss schedulers, leaving
rebuilders for \S\ref{sec-rebuilder}.

Section \S\ref{sec-background} introduced three different \emph{task schedulers}
that decide which tasks to execute and in what order; see the ``Scheduler''
column of Table~\ref{tab-summary} in \S\ref{sec-background-summary}. This
subsection explores the properties of the three schedulers, and possible
implementations.

\subsection{Topological Scheduler}\label{sec-topological}

The topological scheduler pre-computes a linear order of tasks, which when
followed ensures dependencies are satisfied, then executes the required tasks in
that order. Computing such a linear order is straightforward -- given a task
description and output \hs{key} first find the (acyclic) graph of the \hs{key}'s
dependencies, then compute a topological order. Taking the \Make example from
Fig.~\ref{fig-make}, we might compute the order:

\begin{enumerate}
\item \cmd{util.o}
\item \cmd{main.o}
\item \cmd{main.exe}
\end{enumerate}

\noindent
Given the dependencies, we could have equally chosen to build \cmd{main.o}
first, but \cmd{main.exe} \emph{must} come last.

The advantage of this scheme is simplicity -- compute an order, then execute
tasks in that order. In addition any missing nodes or cycles can be detected
from the graph, and reported to the user before any work has commenced.

The downside of this approach is that it requires the dependencies of each task
in advance. As we saw in~\S\ref{sec-deps}, we can only extract dependencies from
an applicative task, which requires the build system to choose
\hs{c}~\hs{=}~\hs{Applicative}, ruling out dynamic dependencies.

\subsection{Restarting Scheduler}\label{sec-restarting}

To handle dynamic dependencies we cannot precompute a static order -- we must
interleave running tasks and ordering tasks. One approach is just to build tasks
in an arbitrary order, but if a task calls \hs{fetch} on an out-of-date key
\cmd{dep}, abort the task and build \cmd{dep} instead. Returning to the example
from Fig.~\ref{fig-make}, we might build in the order:

\begin{enumerate}
\item \cmd{main.exe} (abort because it requires \cmd{util.o})
\item \cmd{util.o}
\item \cmd{main.o}
\item \cmd{main.exe} (restart from scratch, completing successfully this time)
\end{enumerate}

\noindent
We start with \cmd{main.exe} (chosen randomly), but discover it requires
\cmd{util.o}, so instead start building \cmd{util.o}. Next we choose to build
\cmd{main.o} (again, randomly), before finally returning to \cmd{main.exe},
which now has all its dependencies available and completes without error.

The approach works, but has a number of disadvantages. Firstly, it requires a
technical mechanism to abort a task -- easy enough in our theoretical setting of
\hs{Task} (see the implementation in~\S\ref{sec-implementation-excel}) but leads
to engineering concerns in the real world. Secondly, it is not minimal in the
sense that a task may start, do some meaningful work, and then abort, repeating
that same work when restarted.

As a refinement, to reduce the number of aborts (often to zero) \Excel records
the discovered task order in its \emph{calc chain}, and uses it as the starting
point for the next build~(\S\ref{sec-background-excel}). \Bazel's restarting
scheduler does not store the discovered order between build runs; instead, it
stores the most recent task dependency information from which it can compute a
linear order. Since this information may become outdated, \Bazel may also need
to abort a task if a newly discovered dependency is out of date.

\subsection{Suspending Scheduler}\label{sec-suspending}

An alternative approach, utilised by the \hs{busy} build system
(\S\ref{sec-general-build}) and \Shake, is to simply build dependencies when
they are requested, suspending the currently running task. Using the example
from Fig.~\ref{fig-make}, we would build:

\begin{itemize}
\item \cmd{main.exe} (suspended)\\
  $\hookrightarrow$ \cmd{util.o}
\item \cmd{main.exe} (resumed then suspended again)\\
  $\hookrightarrow$   \cmd{main.o}
\item \cmd{main.exe} (completed)
\end{itemize}

\noindent
As we see, the \cmd{main.exe} is started first as it is the required target. It
soon discovers a dependency on \cmd{util.o} and suspends to build \cmd{util.o},
resumes a build before building \cmd{main.o} then continues its work to finish
\cmd{main.exe}.

This scheduler (when combined with a suitable rebuilder) provides a minimal
build system which supports dynamic dependencies. The difficulties are all
engineering in nature. To implement suspension there are two standard
approaches:

\begin{itemize}
\item Blocking threads or processes. This approach is relatively easy, but can
require significant resources, especially if a large number of tasks are
suspended. In languages with cheap green threads (e.g. Haskell) the approach is
more feasible, and it was the original approach taken by \Shake.
\item Continuation-passing style \cite{claessen_continuations} can allow the
remainder of a task to be captured, paused, and resumed later. Continuation
passing is efficient, but requires the build script to be architected to allow
capturing continuations. \Shake currently uses continuations.
\end{itemize}

\noindent
While a suspending scheduler is theoretically optimal, it is only better in
practice than a restarting scheduler if the cost of avoided duplicate work
outweighs the cost of suspending tasks. Note furthermore that the cost of
duplicate work may often be just a fraction of the overall build cost.

\q{Andrey}{It's unclear how our \hs{busy} build system fits into the two above
approaches. It probably is a third approach (i.e. using stack), but it's
unsuitable for real build systems that call external processes. Shall we mention
this?}

\q{Andrey}{A follow-up to the above: our \Shake models use neither of the two
above approaches, which seems a bit unsatisfactory. Can we at least hint how
our models could be adapted to use the continuation-passing style? Can we write
\hs{busyCPS}?}