### 2019-04-28 papers/jfp/10-conclusions.tex
More tidying up of S8

\section{Conclusions}\label{sec-conclusions}

We have investigated multiple build systems, showing how their properties are consequences of two implementation choices: what order you build in and how you decide whether to rebuild. By first decomposing the pieces, we show how to recompose the pieces to find new points in the design space. In particular, a simple recombination leads to a design for a monadic suspending cloud build system, which we have implemented and use in our day-to-day development.