### 2019-04-29 papers/jfp/7-experience.tex
Polish S7-S9

\section{Experience}\label{sec-experience}

As a set of authors, we've written a paper about the \Shake build
system~\cite{mitchell2012shake}, applying \Shake to building GHC~\cite{hadrian},
and a paper about the categorisation of build
systems~\cite{mokhov2018buildsystems}. The original motivation of the research
in that final paper was to create ``\Cloud \Shake'', and apply it to the GHC
build system. In this section we tie these papers together, reflecting on our
experience since these papers.

\subsection{Experience from \Shake}\label{sec-experience-shake}

The original design of \Shake has not changed since the initial paper, but the
implementation has continued to mature -- there have been roughly 5,000
subsequent commits to the \Shake
project\footnote{\url{https://github.com/ndmitchell/shake}}. These commits add
concepts like \emph{resources} (for handling situations when two build tasks
contend on a single external resource), rewriting serialisation to be faster,
documentation including a website\footnote{\url{https://shakebuild.com}}, and
add a lot of tests. The biggest change in that time period was moving from
blocking threads to continuations for the suspending scheduler. The most visible
change is switching from \hs{*>} to \hs{@\%@>} for defining rules\footnote{A
conflicting \hs{*>} operator was added to the Haskell \hs{Prelude}.}, but almost
all external and internal details remain the same.

We consider the lack of change suggestive that \Shake is based on fundamental
principles -- principles we can now name and describe as a consequence
of the categorisation paper~\cite{mokhov2018buildsystems}. There are two main
aspects to the original \Shake paper~\cite{mitchell2012shake} that are described
more clearly in this paper. Firstly, the rebuilder can now be described using
verifying step traces~\S\ref{sec-step-traces}, with a much clearer relationship
to the unoptimised verifying traces of~\S\ref{sec-verifying-traces}. The second
aspect is the power of \Shake tasks, or ``actions'' in the original paper,
which were originally described using this data type:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
data Action k v a = Finished a
                  | Depends k (v -> Action k v a)
\end{minted}
\vspace{1mm}

\noindent
This type is equivalent to \hs{Task}~\hs{Monad} -- the two types form an
isomorphism. The original paper uses concrete types \hs{Key} and \hs{Value}, we
generalise these types to \hs{k} and \hs{v}, and also add \hs{a} so that
\hs{Action}~\hs{k}~\hs{v} can be an instance of \hs{Monad}. We define that
\hs{Monad} instance, and use it to write conversion functions to/from our
\hs{Task} type, witnessing the isomorphism:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
instance Monad (Action k v) where
    return = Finished
    Finished x    >>= f = f x
    Depends ds op >>= f = Depends ds $ \v -> op v >>= f

toAction :: Task Monad k v -> Action k v v
toAction (Task run) = run $ \k -> Depends k Finished

fromAction :: Action k v v -> Task Monad k v
fromAction x = Task $ \fetch -> f fetch x
  where
    f _     (Finished v  ) = return v
    f fetch (Depends d op) = fetch d >>= f fetch . op
\end{minted}
\vspace{1mm}

\noindent
Similarly, \Make tasks were described as equivalent to

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
data Rule k v a = Rule { depends :: [k], action :: [v] -> a }
\end{minted}
\vspace{1mm}

Assuming the lengths of the lists \hs{[@@k]} and \hs{[@@v]} always match, this
type is isomorphic to \hs{Task}~\hs{Applicative}, and we can define a similar
\hs{Applicative} instance and conversion functions. By describing these types
using \hs{Task} we are able to describe the differences more concisely
(\hs{Monad} vs \hs{Applicative}), use existing literature to determine what is
and isn't possible, and explore other constraints beyond just \hs{Monad} and
\hs{Applicative}.

\subsection{Experience from building GHC with \Shake}

\todo{AM}{Finish this section.}

The build system used to compile GHC, known as Hadrian, has continued
development. The use of dynamic dependencies has made the build much easier, and
it's now more maintainable. It's now merged into the GHC repo and tested as
standard by the CI. It's current state is ... In particular, it takes N minutes
to build with \Shake, vs M minutes to build with \Make. The \Shake version has
attracted more contributors, and been easier to modify, etc. The code is N
lines.

The design we originally outlined remains the one in use. There has been a huge
amount of engineering work to make the tests pass (many of the details encoded
in the \Make build system were incidental -- but some were essential).

One of the benefits of using Shake is that we have access to high quality
profiling information, allowing us to compute critical paths and other metrics,
see Mitchell~\shortcite{mitchell2019ghcrebuildtimes} for a tour of the metrics.
This information has shown us that GHC is slow to build (40m on an 8 CPU machine
on Windows), that more CPUs would not help (on unlimited CPUs it would take at
least 37m), and that a handful of steps (two Haskell compiles, some calls to
\cmd{configure}) are responsible for a significant amount of that time (at least
10m).

\subsection{Experience from \Cloud \Shake}\label{sec-cloud-shake}

Converting \Shake into \Cloud \Shake was not a difficult process once armed with
a roadmap~\cite{mokhov2018buildsystems}. The key was the introduction of two new
functions:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
addCloud :: k -> Ver -> Ver -> [[(k, Hash v)]]
         -> v -> [FilePath] -> IO ()
lookupCloud :: (k -> m (Maybe (Hash v))) -> k -> Ver -> Ver
            -> m (Maybe (v, [[k]], IO ()))
\end{minted}
\vspace{1mm}

\noindent
These functions are suspiciously like \hs{recordCT} and \hs{constructCT}
from~\S\ref{sec-constructive-traces}, with their differences perhaps the most
illustrative of the changes required. (We have made some minor changes from
actual \Shake, like replacing \hs{Key} for \hs{k}, to reduce irrelevant
differences.)

\begin{itemize}
\item There are two \hs{Ver} arguments being passed to each function. These are
      the versions of the build script, and the rule for this particular key. If
      either version changes then it is as though the key has changed, and
      nothing will match. These versions are important to avoid using stale
      build products from previous versions of the build script.
\item The list of dependencies to \hs{addCloud} is a list of lists, rather than
      a simple list. The reason is that \Shake allows a list of dependencies to
      be supplied at once, so they can all be built in parallel.
\item The \hs{addCloud} function also takes a list of \hs{FilePath}, being the
      files that this rule produces. These produced files include those which
      are output keys from a rule and those declared with the function
      \hs{produces}.
\item The \hs{lookupCloud} function allows an explicit \hs{Nothing} when looking
      up a dependent key, since some keys are not buildable.
\item The \hs{lookupCloud} function returns at most one result, rather than a
      list. This change was made for simplicity.
\end{itemize}

To integrate these functions into \Shake we found the most expedient route was
to leave \Shake with verifying traces, but if the verifying trace does not
match, we consult the constructive trace. By bolting constructive traces onto
the side of Shake we avoid re-engineering of the central database. We have not
found any significant downsides from the bolt-on approach thus far, so it may be
a sensible route to go even if developing from scratch -- allowing an optimised
verified trace implementation in many cases, and falling back to a more complex
implementation (requiring consulting remote servers) only rarely.

The one thing we have not yet completed on the engineering side is a move to
hosting caches over HTTP. At the moment all caches are on shared file systems.
This approach can use mounted drives to mirror HTTP connections onto file
systems, and reuse tools for managing file systems, share caches with
\cmd{rsync}, and is simple. Unfortunately, on certain operating systems (e.g.
Windows) mounting an HTTP endpoint as a file system requires administrator
privileges, so an HTTP cache is still desirable.

\subsection{Experience from using \Cloud \Shake}\label{sec-using-cloud-shake}

While we expected the GHC build system to be the first to take advantage of
\Cloud \Shake, we were actually beaten to it by Standard Chartered who
report\footnote{\url{https://groups.google.com/d/msg/shake-build-system/NbB5kMFS34I/mZ9L4TgkBwAJ}}:

\vspace{1mm}
\begin{center}
\parbox{0.8 \textwidth}{\emph{Thanks for the symlinks release, we just finished upgrading this build
system to use \cmd{-}\cmd{-share}. ... Building from scratch with a warm cache
takes around 5 seconds, saving us up to 2 hours. Not bad!}}
\end{center}
\vspace{1mm}

We have also got the GHC build working with \Cloud \Shake, resulting in build
times of 3m and 30s instead of 1h and 8m (on Windows). Due to aspects like slow
and fully sequential \cmd{configure} we do not see the dramatic reductions
reported by Standard Chartered. Converting to a build suitable for sharing is
not overly onerous, but nor is it trivial. Some things that require attention:

\begin{itemize}
\item \textbf{Irrelevant differences}: A common problem is that you do not get
      shared caching when you want it. To get caching, it is important that all
      machines are identical for the aspects that involve the build system. In
      particular, that means that all paths must be relative, and all
      dependencies on system binaries (e.g. \cmd{gcc}) should either be
      controlled by the build system (e.g. using \Nix) or have tolerance to
      account for minor machine variations.
\item \textbf{Insufficient produced files}: A build rule must declare all files
      it produces, so these are stored in the cache and can be restored from the
      cache. If you do not use a remote cache then it is permissible to have
      files not known to the build system which change at the same time as files
      that are known. Once using a remote cache any file that is produced as a
      side-effect of a rule and then later used \emph{must} be declared. Failure
      to declare such files means that if some step produces a file, and a later
      step relies on it, that step breaks. In practice, most issues encountered
      during the move to cloud builds for GHC were a lack of declared produced
      files.
\item \textbf{Insufficient dependencies}: While insufficient dependencies are
      often a problem, the move to a cloud build makes them even more critical,
      as missing dependencies are likely to mean a result is \emph{never built
      locally}, rather than just build locally only once. Furthermore,
      dependencies in a cloud setting must be more complete. For example, using
      the build system from Fig.~\ref{fig-make}, with a local build system
      without early cutoff, if the rule to produce \cmd{main.exe} also made use
      of \cmd{main.c} but did not declare so, that is not a problem. With the
      addition of early cutoff, if \cmd{main.c} changes in an important way but
      \cmd{main.o} is not rebuilt that causes issues, but is a somewhat rare
      occurrence. With cloud builds that becomes somewhat more common.
\end{itemize}

To help with the final two issues -- insufficient dependencies and produced
files -- we have further enhanced the \Shake lint modes, coupling them to a
program called \cmd{FSATrace}, which detects which files are read/written by a
command line execution. Such information has been very helpful in making the GHC
build cloud ready.