### 2019-04-29 papers/jfp/9-related.tex

\section{Related work}\label{sec-related}

While there is research on individual build systems, there has been little
research to date comparing different build systems. In~\S\ref{sec-background} we
covered several important build systems~--~in this section we relate a few
other build systems to our abstractions, and discuss other work where similar
abstractions~arise.

\subsection{Other Build Systems}\label{sec-related-build}

Most build systems, when viewed at the level we talk, can be captured with minor
variations on the code presented in \S\ref{sec-implementations}. Below we list
some notable examples:

\begin{itemize}
\item \Dune~\cite{dune} is a build system designed for OCaml/Reason projects.
Its distinguishing feature is that it uses
\emph{arrows}~\cite{hughes2000generalising} rather than monads to model
dynamic dependencies, which simplifies static dependency approximation.

\item \Ninja~\cite{ninja} combines the \hs{topological} scheduler of \Make with
the verifying traces of \Shake~--~our associated implementation provides such a
combination. \Ninja~is also capable of modelling build rules that produce
multiple results, a limited form of polymorphism \S\ref{sec-polymorphism}.

\item \Nix~\cite{dolstra2004nix} has coarse-grained dependencies, with precise
hashing of dependencies and downloading of precomputed build products. We
provided a model of \Nix in \S\ref{sec-implementation-cloud}, although it is
worth noting that \Nix is not primarily intended as a build system, and the
coarse grained nature (packages, not individual files) makes it targeted to a
different purpose.

\item \Pluto~\cite{erdweg2015pluto} is based on a similar model to \Shake, but
additionally allows cyclic build rules combined with a user-specific resolution
strategy. Often such a strategy can be unfolded into the user rules without loss
of precision, but a fully general resolution handler extends the \hs{Task}
abstraction with additional features.

\item \Redo~\cite{redo-idea,grosskurth2007redo,redo} almost exactly
matches \Shake at the level of detail given here, differing only on aspects like
polymorphic dependencies~\S\ref{sec-polymorphism}.

\item \Tup~\cite{tup} functions much like \Make, but with a refined dirty-bit
implementation that watches the file system for changes and can thus avoid
rechecking the entire graph. \Tup also automatically deletes stale results.
\end{itemize}

The one build system we are aware of that cannot be modelled in our framework is
\Fabricate by~Hoyt~\etal~\shortcite{fabricate}. In \Fabricate a build system is
a script that is run in-order, in the spirit of:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{bash}
gcc -c util.c
gcc -c main.c
gcc util.o main.o -o main.exe
\end{minted}
\vspace{1mm}

\noindent
To achieve minimality, each separate command is traced at the OS-level, allowing
\Fabricate to record a trace entry stating that \cmd{gcc -c util.c} reads from
\cmd{util.c}. In future runs \Fabricate runs the script from start to finish,
skipping any commands where no inputs have changed. The key difference from our
\hs{Tasks} abstraction is that instead of supplying a mapping from outputs to
tasks, \Fabricate supplies a list of statements, in an order, without declaring
what each line produces. There is no need to schedule the statements, and not
enough information to do so.

Taking our abstraction, it is possible to encode \Fabricate assuming that
commands like \cmd{gcc -c util.c} are keys, there is a linear dependency between
each successive key, and that the OS-level tracing can be lifted back as a
monadic \hs{Task} function\footnote{\Shake provides support for
\Fabricate{}-like build systems~--~see \cmd{Development.Shake.Forward}.}.
However, in our pure model the mapping is not perfect as \cmd{gcc} writes to
arbitrary files whose locations are not known in advance.

\subsection{Self-adjusting Computation}

While not typically considered build systems, self-adjusting computation is a
well studied area, and in particular the contrast between different formulations
has been thoroughly investigated, e.g.
see~Acar~\etal~\shortcite{acar2007selfadjusting}. Self-adjusting computations
can automatically adjust to an external change to their inputs. A classic
example is a self-adjusting sorting algorithm, which can efficiently (in
$O(\log{n})$ time where $n$ is the length of the input) recalculate the result
given an incremental change of the input. While very close to build systems in
spirit, self-adjusting computations are mostly used for in-memory computation
and rely on the ability to dynamically allocate new keys in the store for
sharing intermediate computations~--~an intriguing feature rarely seen in build
systems (\Shake's oracles~\S\ref{sec-polymorphism} can be used to model this
feature to a limited degree).

A lot of research has been dedicated to finding efficient data structures and
algorithms for self-adjusting computations (with a few open-source
implementations, e.g. \Incremental by Jane Street~\shortcite{incremental}. We
plan to investigate how these insights can be utilised by build systems as
future work.

\subsection{Memoization}\label{sec-related-memo}

\emph{Memoization} is a classic optimisation technique for storing values of a
function instead of recomputing them each time the function is called. Minimal
build systems (see the Definition~\ref{def-minimal}) certainly perform
memoization: they \emph{store values instead of recomputing them each time}.
Memoization can therefore be reduced to a minimal build system (as we
demonstrate below), but not vice versa, since minimal build systems solve a more
complex optimisation problem.

As a simple example of using a build system for memoization, we solve a textbook
dynamic programming problem~--~Levenshtein's \emph{edit
distance}~\cite{levenshtein1966binary}: given two input strings $a$ and
$b$, find the shortest series of edit operations that transforms $a$
to $b$. The edit operations are typically \emph{inserting}, \emph{deleting} or
\emph{replacing} a symbol. The dynamic programming solution of this problem is
so widely known, e.g., see~Cormen~\etal~\shortcite{cormen2001introduction}, that
we provide its encoding in our \hs{Tasks} abstraction without further
explanation. We address elements of strings $a_i$ and $b_i$ by keys \hs{A}~$i$
and \hs{B}~$i$, respectively, while the cost of a subproblem $c_{ij}$ is
identified by \hs{C}~$i$~$j$.

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
data Key = A Int | B Int | C Int Int deriving Eq
\end{minted}
\vspace{0mm}
\begin{minted}[xleftmargin=10pt]{haskell}
editDistance :: Tasks Monad Key Int
editDistance (C i 0) = Just $ Task $ const $ pure i
editDistance (C 0 j) = Just $ Task $ const $ pure j
editDistance (C i j) = Just $ Task $ \fetch -> do
    ai <- fetch (A i)
    bj <- fetch (B j)
    if ai == bj
        then fetch (C (i - 1) (j - 1))
        else do insert  <- fetch (C  i      (j - 1))
                delete  <- fetch (C (i - 1)  j     )
                replace <- fetch (C (i - 1) (j - 1))
                return (1 + minimum [insert, delete, replace])
editDistance _ = Nothing
\end{minted}
\vspace{1mm}

\noindent
When asked to build \hs{C}~$n$~$m$, a minimal build system will calculate the
result using memoization. Furthermore, when an input symbol $a_i$ is changed,
only necessary, incremental recomputation will be performed~--~an optimisation
that cannot be achieved just with memoization.

Self-adjusting computation, memoization and build systems are inherently related
topics, which poses the question of whether there is an underlying common
abstraction waiting to be discovered.