### 2019-05-03 papers/jfp/10-conclusions.tex
Restore acknowledgements

\section{Conclusions}\label{sec-conclusions}

We have investigated multiple build systems, showing how their properties are
consequences of two implementation choices: what order you build in and how you
decide whether to rebuild. By first decomposing the pieces, we show how to
recompose the pieces to find new points in the design space. In particular, a
simple recombination leads to a design for a monadic suspending cloud build
system, which we have implemented and use in our day-to-day development.

\section*{Acknowledgements}

Thanks to anonymous reviewers and everyone else who provided us with feedback
on earlier drafts: Ulf Adams, Arseniy Alekseyev, Dan Bentley, Martin
Br\"{u}stel, Ulan Degenbaev, Jeremie Dimino, Andrew Fitzgibbon, Georgy
Lukyanov, Simon Marlow, Evan Martin, Yaron Minsky, Guillaume Maudoux, Philip
Patsch, Michael Peyton Jones, Andrew Phillips, Fran\c{c}ois Pottier, Rohit
Ramesh, Irakli Safareli, Zhen Zhang. Your contributions were incredibly
valuable.

Andrey Mokhov is funded by a Royal Society Industry Fellowship \cmd{IF160117}
on the topic ``Towards Cloud Build Systems with Dynamic Dependency Graphs''.