### 2019-10-17 papers/jfp/2-background.tex

\section{Background}\label{sec-background}

Build systems automate the execution of repeatable tasks, at a scale from individual
users up to large organisations. In this section we explore the design space of build
systems, using four examples: \Make~\cite{feldman1979make},
\Shake~\cite{mitchell2012shake}, \mbox{\Bazel~\cite{bazel}}, and
\Excel~\cite{advanced_excel}.
We have carefully chosen these four to illustrate the various axes on which
build systems differ; we discuss many other notable examples of build systems,
and their relationships, in~\S\ref{sec-implementations} and~\S\ref{sec-related}.

\subsection{The Venerable \Make: Static Dependencies and File Modification Times}
\label{sec-background-make}

\Make\footnote{There are numerous implementations of \Make and none comes with a
formal specification. In this paper we use a simple approximation to a real
\Make that you might find on your machine.} was developed
more than 40 years ago to automatically build software libraries and executable
programs from source code. It uses \emph{makefiles} to describe \emph{tasks} --- often
referred to as \emph{build rules} --- and their \emph{dependencies}, in a simple textual
form. For example:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{makefile}
util.o: util.h util.c
    gcc -c util.c
\end{minted}
% \vspace{1mm}
\begin{minted}[xleftmargin=10pt]{makefile}
main.o: util.h main.c
    gcc -c main.c
\end{minted}
% \vspace{1mm}
\begin{minted}[xleftmargin=10pt]{makefile}
main.exe: util.o main.o
    gcc util.o main.o -o main.exe
\end{minted}
\vspace{1mm}

\noindent
The above makefile lists three tasks: (i) compile a utility library comprising
files \cmd{util.h} and \cmd{util.c} into \cmd{util.o} by
executing\footnote{In this example we pretend \cmd{gcc} is a pure function for
  the sake of simplicity. In reality, there are multiple versions of \cmd{gcc}.
  To account for this, the actual binary for \cmd{gcc} is often listed as a dependency, along
  with any supporting binaries, or standard libraries (such as \cmd{stdio.h}),
  that are used by \cmd{gcc}.
} the command
\cmd{gcc -c util.c}, (ii) compile the main source file \cmd{main.c} into
\cmd{main.o}, and (iii) link object files \cmd{util.o} and \cmd{main.o} into the
executable \cmd{main.exe}. The makefile contains the complete information about
the \emph{task dependency graph}, which is shown in Fig.~\ref{fig-make}(a).

\begin{figure}%[h]
\begin{subfigure}[b]{0.32\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/make-example.pdf}}
\caption{Task dependency graph}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.32\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/make-example-full.pdf}}
\caption{Full rebuild}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.32\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/make-example-partial.pdf}}
\caption{Partial rebuild}
\end{subfigure}
% \vspace{-2mm}
\caption{A task dependency graph and two build scenarios. Input files are shown
in rectangles, intermediate and output files are shown in rounded rectangles.
Modified inputs and files that are rebuilt are highlighted.
\label{fig-make}}
% \vspace{-2mm}
\end{figure}

If the user runs \Make specifying \cmd{main.exe} as the desired output, \Make
will build~\cmd{util.o} and~\cmd{main.o}, in any order (or even in parallel) since these tasks are
independent, and then \cmd{main.exe}. If the user modifies \cmd{util.h} and runs
\Make again, it will perform a \emph{full rebuild}, because all three tasks
transitively depend on \cmd{util.h}, as illustrated in Fig.~\ref{fig-make}(b).
On the other hand, if the user modifies \cmd{main.c} then a \emph{partial
rebuild} is sufficient: \cmd{util.o} does not need to be rebuilt, since its
inputs have not changed, see Fig.~\ref{fig-make}(c). Note that if the dependency
graph is \emph{acyclic} then each task needs to be executed at most once. Cyclic
task dependencies are typically not allowed in build systems, although there are rare
exceptions, see~\S\ref{sec-iterative-compute}.

\noindent
The following property is essential for build systems; indeed, it is their
raison d'\^etre:

\vspace{2mm}
\textbf{Definition} (Minimality). A build system is \emph{minimal} if it
executes tasks at most once per build, and only if they transitively depend on
inputs that changed since the previous build.\label{def-minimal}
\vspace{2mm}

To achieve minimality \Make relies on two main ideas: (i) it uses \emph{file
modification times} to detect which files changed\footnote{Technically, you
can fool \Make by altering the modification time of a file without changing its
content, e.g. by using the command \cmd{touch}. \Make is therefore minimal only
under the assumption that you do not do that.}, and (ii) it constructs a task
dependency graph from the information contained in the makefile and executes
tasks in a \emph{topological order}. For a more concrete description
see~\S\ref{sec-topological} and~\S\ref{sec-implementation-make}.

\subsection{\Excel: Dynamic Dependencies at the Cost of Minimality}
\label{sec-background-excel}

\Excel is a build system in disguise. Consider the following simple spreadsheet.

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{text}
A1: 10      B1: A1 + A2
A2: 20
\end{minted}
\vspace{1mm}

\noindent
There are two input cells \cmd{A1} and \cmd{A2}, and a single task that computes
the sum of their values, producing the result in cell \cmd{B1}. If either of
the inputs change, \Excel will recompute \cmd{B1}.

Unlike \Make, \Excel does not need to know all task dependencies upfront.
Indeed, some dependencies may change \emph{dynamically} during computation. For
example:

\vspace{0.5mm}
\begin{minted}[xleftmargin=10pt]{text}
A1: 10      B1: INDIRECT("A" & C1)      C1: 1
A2: 20
\end{minted}
\vspace{0.5mm}

\noindent
The formula in \cmd{B1} uses the \cmd{INDIRECT} function, which takes a string
and returns the value of the cell with that name.  The string evaluates to
\cmd{"A1"}, so \cmd{B1} evaluates to \cmd{10}. However, the dependencies of the
formula in \cmd{B1} are determined by the value of \cmd{C1}, so it is impossible
to compute the dependency graph before the build starts. In this particular
example the value of \cmd{C1} is a constant, but
it might instead be the result of a long computation chain --
so its value will only
become available during the build.

To support dynamic dependencies, \Excel's calculation engine~\cite{excel_recalc}
is significantly different from \Make. \Excel arranges the cells into a linear
sequence, called the \emph{calc chain}.  During the build, \Excel processes
cells in the calc-chain sequence, but if computing a cell \cmd{C} requires the
value of a cell \cmd{D} that has not yet been computed, \Excel \emph{aborts}
computation of \cmd{C}, moves \cmd{D} before \cmd{C} in the calc chain, and
resumes the build starting with \cmd{D}. When a build is complete, the resulting
calc chain respects all the dynamic dependencies of the spreadsheet. When an
input value or formula is changed, \Excel uses the final calc chain from the
\emph{previous} build as its starting point so that, in the common case where
changing an input value does not change dependencies, there are no aborts.
Notice that build always succeeds regardless of the initial calc chain (barring
truly circular dependencies); the calc chain is just an optimisation.
We refer to this algorithm as \emph{restarting}, and discuss it
in more detail in~\S\ref{sec-restarting} and~\S\ref{sec-implementation-excel}.

Dynamic dependencies complicate minimality.  In the above example, \cmd{B1}
should only be recomputed if \cmd{A1} or \cmd{C1} change, but not if (say)
\cmd{A2} changes; but these facts are not statically apparent. In practice
\Excel implements a conservative approximation to minimality: it recomputes a
formula if~(i)~the formula statically mentions a changed cell, or~(ii)~the formula uses a
function like \cmd{INDIRECT} whose dependencies are not statically visible,
or~(iii)~the formula itself has changed.

Item (iii) in the above list highlights another distinguishing feature of \Excel: it
is \emph{self-tracking}. Most build systems only track changes of inputs and
intermediate results, but \Excel also tracks changes in the tasks themselves: if
a formula is modified, \Excel will recompute it and propagate the changes.
Self-tracking is uncommon in software build systems, where one often needs to
manually initiate a full rebuild even if just a single task has changed.
We discuss self-tracking further in~\S\ref{sec-tracking-aspects}.

\begin{figure}
\begin{subfigure}[b]{0.90\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/shake-example.pdf}}
\caption{Dependency graph produced after the previous build.}
\end{subfigure}
\begin{subfigure}[b]{0.90\linewidth}
\vspace{1mm}
\centerline{\includegraphics[scale=0.28]{fig/shake-example-rebuild.pdf}}
\caption{The input file \cmd{docs.txt} was modified, hence we rebuild
\cmd{release.txt} and \cmd{release.tar}, discovering a new dependency
\cmd{README} in the process.}
\end{subfigure}
% \vspace{-2mm}
\caption{Dynamic dependencies example: create \cmd{README} and add it to the
list of release documents \cmd{docs.txt}.\label{fig-shake}}
% \vspace{-4mm}
\end{figure}

\vspace{-1mm}
\subsection{\Shake: Dynamic Dependencies with No Remorse}
\label{sec-background-shake}

\Shake was developed to solve the issue of dynamic
dependencies~\cite{mitchell2012shake} without sacrificing
the minimality requirement. Building on the \Make example
from~\S\ref{sec-background-make}, we add the following files whose
dependencies are shown in Fig.~\ref{fig-shake}(a):

\begin{itemize}
    \item \cmd{LICENSE} is an input text file containing the project license.
    \item \cmd{release.txt} lists all release files. This file is produced by
    concatenating input text files \cmd{bins.txt} and \cmd{docs.txt}, which list
    binary and documentation files of the project.
    \item \cmd{release.tar} is the archive built by executing the command
    \cmd{tar} on the release files.
\end{itemize}

The dependencies of \cmd{release.tar} are not known statically: they are
determined by the content of \cmd{release.txt}, which might not even exist
before the build. Makefiles cannot express such dependencies, requiring
workarounds such as \emph{build phases}, which are known to be
problematic~\cite{hadrian}. In \Shake we can express the rule for
\cmd{release.tar} as:

\begin{minted}[xleftmargin=10pt]{haskell}
"release.tar" %> \_ -> do
    need ["release.txt"]
    files <- lines <$> readFile "release.txt"
    need files
    cmd "tar" $ ["-cf", "release.tar"] ++ files
\end{minted}

\noindent
We first declare the static dependency on \cmd{release.txt}, then read its
content (a list of files) and depend on each listed file, dynamically. Finally,
we specify the
command to produce the resulting archive. Crucially, the archive will only be
rebuilt if one of the dependencies (static or dynamic) has changed. For example,
if we create another documentation file \cmd{README} and add it to
\cmd{docs.txt}, \Shake will appropriately rebuild \cmd{release.txt} and
\cmd{release.tar}, discovering the new dependency, see Fig.~\ref{fig-shake}(b).

\Shake's implementation is different from both \Make and \Excel in two aspects.
First, to decide which files need to be rebuilt,
it uses the dependency graph that is dynamically constructed during
the previous build. This idea has a long history, going back to
\emph{incremental}~\cite{demers1981incremental},
\emph{adaptive}~\cite{acar2002adaptive}, as well as
\emph{self-adjusting computations} -- see
Acar~\etal~\shortcite{acar2007selfadjusting} and~\S\ref{sec-related}. Second,
instead of aborting and deferring the execution of tasks whose newly discovered
dependencies have not yet been built (as \Excel does), \Shake \emph{suspends}
their execution until the dependencies are brought up to date. We refer to this
task scheduling algorithm as \emph{suspending}, see a further discussion
in~\S\ref{sec-suspending} and a concrete implementation
in~\S\ref{sec-implementation-shake}.

\begin{figure}
% \vspace{-3mm}
\centerline{\includegraphics[scale=0.28]{fig/shake-example-cutoff.pdf}}
% \vspace{-4mm}
\caption{An early cutoff example: if a comment is added to \cmd{main.c}, the
rebuild is stopped after detecting that \cmd{main.o} is unchanged, since this
indicates that \cmd{main.exe} and its dependents do not need to be
rebuilt.\label{fig-cutoff}}
% \vspace{-4mm}
\end{figure}

\Shake also supports the \emph{early cutoff optimisation}. When it
executes a task and the result is unchanged from the previous build, it is
unnecessary to execute the dependent tasks, and hence \Shake can stop a build
earlier, as illustrated in Fig.~\ref{fig-cutoff}. Not all build systems support
early cutoff: \Shake and \Bazel (introduced below) do, but \Make and \Excel do
not; see~\S\ref{sec-dirty-bit} for an explanation of why.

% \vspace{-0.5mm}
\subsection{\Bazel: A Cloud Build System}\label{sec-background-bazel}
% \vspace{-0.5mm}

When build systems are used by large teams, different team members often end up
executing exactly the same tasks on their local machines. A \emph{cloud build
system} can speed up builds dramatically by sharing build results
among team members. Furthermore, cloud build systems can support \emph{shallow
builds} that materialise only end build products locally, leaving all
intermediates in the cloud.

\begin{figure}
% \vspace{-2mm}
\begin{subfigure}[b]{0.25\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/bazel-example-checkout.pdf}}
\vspace{-0.5mm}
\caption{Download sources}
\end{subfigure}
\begin{subfigure}[b]{0.40\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/bazel-example-build.pdf}}
\vspace{-0.5mm}
\caption{Build \cmd{main.exe}}
\end{subfigure}
\begin{subfigure}[b]{0.31\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/bazel-example-rebuild.pdf}}
\vspace{-0.5mm}
\caption{Modify \cmd{util.c} and rebuild}
\end{subfigure}
\vspace{-1mm}
\caption{A cloud build example: (a)~download sources, (b)~build \cmd{main.exe}
by downloading it from the cloud and skipping intermediate files (only their
hashes are needed), (c)~modify \cmd{util.c} and rebuild \cmd{main.exe}, which
requires building \cmd{util.o} (nobody has compiled \cmd{util.c} before) and
downloading \cmd{main.o} (it is needed for linking \cmd{main.exe}). File hashes
are shown in circles, and non-materialised intermediates in dashed rounded
rectangles.\label{fig-bazel}}
\vspace{-1mm}
\end{figure}

Consider the example in Fig.~\ref{fig-bazel}. The user starts by downloading the
sources, whose content hashes are (for simplicity)~1,~2 and~3, and requests to
build \cmd{main.exe}, see Fig.~\ref{fig-bazel}(a,b). By looking up the global
history of all previous builds\footnote{In practice, old entries are regularly
evicted from the cloud storage, as further discussed
in~\S\ref{sec-cloud-aspects}.}, the build system finds that someone has already
compiled these exact sources before and the resulting files \cmd{util.o} and
\cmd{main.o} had hashes 4~and~5, respectively. Similarly, the build system finds
that the hash of the resulting \cmd{main.exe} was 6 and downloads the actual
binary from the cloud storage -- it must be materialised, because it is the end
build product.

In Fig.~\ref{fig-bazel}(c), the user modifies the source file \cmd{util.c},
thereby changing its hash from~1~to~7. The cloud lookup of the new combination
$\{\cmd{util.c}, \cmd{util.h}\}$ fails, which means that nobody has ever
compiled it. The build system must therefore build \cmd{util.o}, materialising
it with the new hash~8. The combination of hashes of \cmd{util.o} and
\cmd{main.o} has not been encountered before either, thus the build system first
downloads \cmd{main.o} from the cloud and then builds \cmd{main.exe} by linking
the two object files. When the build is complete, the results can be uploaded to
the cloud for future reuse by other team members.

\Bazel is one of the first openly-available cloud build systems.
As of writing, it is not possible to express dynamic dependencies in
user-defined build rules; however some of the pre-defined build rules require
dynamic dependencies and the internal build engine can cope with them by using
a \emph{restarting} task scheduler, which is similar to that of \Excel but does
not use the calc chain. \Bazel is not minimal in the sense that it may restart a
task multiple times as new dependencies are discovered and rebuilt, but it
supports the early cutoff optimisation.

To support cloud builds, \Bazel maintains (i)~a \emph{content-addressable cache}
that maps the hash of a file's content to the actual content of that file;
(ii)~a memo table that records all executed build commands
with their input and output file hashes.
The memo table allows the build engine to bypass the execution
of a task, by predicting the hash of the result from the hashes of its
dependencies; then the content-addressable cache allows the engine
to download the result (if needed) based on the result hash. A concrete
implementation is provided in~\S\ref{sec-implementation-cloud}.

% \vspace{-1.5mm}
\subsection{Summary}\label{sec-background-summary}
% \vspace{-0.5mm}

We summarise differences between four discussed build systems in
Table~\ref{tab-summary}. The column \emph{`persistent build information'} refers
to the information that build systems persistently store between builds:
\begin{itemize}
    \item \Make stores file modification times, or rather, it relies on the file
    system to do that.
    \item \Excel stores one dirty bit per cell and the calc chain from the
    previous build.
    \item \Shake stores the dependency graph discovered in the previous build,
    annotated with file content hashes for efficient checking of file changes.
    \item \Bazel stores the content-addressable cache and the history of all
    previous build commands annotated with file hashes. This information is
    shared among all users.
\end{itemize}

In this paper we elucidate which build system properties are consequences of
specific implementation choices (stored metadata and task scheduling algorithm),
and how one can obtain new build systems with desired properties by recombining
parts of existing implementations. As a compelling example,
in~\S\ref{sec-implementation-cloud} we demonstrate how to combine the advantages
of \Shake and \Bazel.

\begin{table}
% \vspace{-2mm}
\caption{Build system differences.\label{tab-summary}}
% \vspace{-4mm}
% \smaller
\centering
\begin{tabular}{llllccc}
\hline
Build system\gap & Persistent build information\gap & Scheduler\gap   & Dependencies\gap    & Minimal\gap & Cutoff\gap & Cloud\\\hline
\Make       & File modification times      & Topological\gap & Static          & Yes     & No     & No   \\
\Excel      & Dirty cells, calc chain      & Restarting  & Dynamic         & No      & No     & No   \\
\Shake      & Previous dependency graph    & Suspending  & Dynamic         & Yes     & Yes    & No   \\
\Bazel      & Cloud cache, command history\gap & Restarting  & Dynamic$^{(*)}$ & No      & Yes    & Yes  \\\hline
\end{tabular}
% \vspace{2mm}
$^{(*)}$At present, user-defined build rules cannot have dynamic dependencies.
% \vspace{-8mm}
\end{table}