### 2019-10-19 papers/jfp/3-abstractions.tex
Polish on S3-4

\section{Build systems, abstractly}\label{sec-abstractions}
% \vspace{-0.5mm}

We have introduced a number of components and
characteristics of build systems:
tasks, dependencies, early cutoff, minimality, etc.
It is easy to get confused.
To make all this more concrete, this section presents executable
abstractions that can express
all the intricacies of build systems discussed in~\S\ref{sec-background},
and allow us to construct complex build systems from simple primitives. Specifically, we
present the \emph{task} and \emph{build} abstractions in~\S\ref{sec-task}
and~\S\ref{sec-general-build}, respectively. Sections~\S\ref{sec-scheduler},
\S\ref{sec-rebuilder} and~\S\ref{sec-implementations} scrutinise the
abstractions further and provide concrete implementations for several build
systems.

% \vspace{-1.5mm}
\subsection{Common Vocabulary for Build Systems}\label{sec-vocabulary}
% \vspace{-0.5mm}

\emph{Keys, values, and the store.} The goal of any build system is to
bring up to date a \emph{store} that implements a mapping from \emph{keys} to
\emph{values}. In software build systems the store is the file system, the
keys are filenames, and the values are file contents. In \Excel, the store is
the worksheets, the keys are cell names (such as \cmd{A1}) and the values are
numbers, strings, etc., displayed as the cell contents. Many build systems use
\emph{hashes} of values as compact summaries with a fast equality check.

\emph{Input, output, and intermediate values.} Some values must be provided by
the user as \emph{input}. For example, \cmd{main.c} can be edited by the user
who relies on the build system to compile it into \cmd{main.o} and subsequently
\cmd{main.exe}. End build products, such as \cmd{main.exe}, are \emph{output}
values. All other values (in this case \cmd{main.o}) are \emph{intermediate};
they are not interesting for the user but are produced in the process of turning
inputs into outputs.

\emph{Persistent build information.} As well as the key/value mapping, the
store also contains information maintained by the build system itself, which
persists from one invocation of the build system to the next -- its ``memory''.

\emph{Task description.} Any build system requires the user to specify how
to compute the new value for one key, using the (up to date) values of its
dependencies. We call this specification the \emph{task description}. For
example, in \Excel, the formulae of the spreadsheet constitute the task
description; in \Make the rules in the makefile are the task description.

\emph{Build system.} A \emph{build system} takes a task description, a
\emph{target key}, and a store, and returns a new store in which the target key
and all its dependencies have up to date values.

We model a build system concretely, as a Haskell program. To that end,
Fig.~\ref{fig-types} provides the type signatures for all key abstractions
introduced in the paper. For example, \hs{Store}~\hs{i}~\hs{k}~\hs{v} is the
type of stores, with several associated functions (\hs{getValue}, etc.). We use
\hs{k} as a type variable ranging over keys, \hs{v} for values, and \hs{i} for
the persistent build information. Fig.~\ref{fig-stdlib} lists standard library
definitions.

\begin{figure}
% \vspace{-1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Abstract store containing a key/value map and persistent build information
data Store i k v -- i = info, k = key, v = value
initialise :: i -> (k -> v) -> Store i k v
getInfo    :: Store i k v -> i
putInfo    :: i -> Store i k v -> Store i k v
getValue   :: k -> Store i k v -> v
putValue   :: Eq k => k -> v -> Store i k v -> Store i k v
\end{minted}
% \vspace{-1mm}
\begin{minted}[fontsize=\small]{haskell}
data Hash v -- a compact summary of a value with a fast equality check
hash    :: Hashable v => v -> Hash v
getHash :: Hashable v => k -> Store i k v -> Hash v
\end{minted}
% \vspace{-1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Build tasks (see §3.2)
newtype Task  c k v = Task (@\std{forall}@ f. c f => (k -> f v) -> f v)
type    Tasks c k v = k -> Maybe (Task c k v)
\end{minted}
% \vspace{-1mm}
\begin{minted}[fontsize=\small]{haskell}
run :: c f => Task c k v -> (k -> f v) -> f v
run (Task task) fetch = task fetch
\end{minted}
% \vspace{-1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Build system (see §3.3)
type Build c i k v = Tasks c k v -> k -> Store i k v -> Store i k v
\end{minted}
% \vspace{-1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Build system components: a scheduler (see §4) and a rebuilder (see §5)
type Scheduler c i ir k v = Rebuilder c ir k v -> Build c i k v
type Rebuilder c   ir k v = k -> v -> Task c k v -> Task (MonadState ir) k v
\end{minted}
% \vspace{-3.5mm}
\caption{Type signatures of key build systems abstractions.}\label{fig-types}
% \vspace{-4.5mm}
\end{figure}

\subsection{The Task Abstraction}\label{sec-task}

Our first main abstraction is for \emph{task descriptions}:
\begin{minted}[xleftmargin=10pt]{haskell}
newtype Task  c k v = Task (@\std{forall}@ f. c f => (k -> f v) -> f v)
type    Tasks c k v = k -> Maybe (Task c k v)
\end{minted}

\noindent
Here \hs{c} stands for \emph{constraint}, such as \hs{Applicative}
(\S\ref{sec-why-polymorphism} explains why we need it).
A \hs{Task} describes a single build task, while \hs{Tasks} associates a
\hs{Task} with every non-input key; input keys are associated with \hs{Nothing}.
The highly-abstracted type \hs{Task} describes how to build a value given
a way to build its dependencies, and is best explained by an example.
Consider this \Excel spreadsheet:
% \footnote{Readers familiar with \emph{lenses} or \emph{profunctor optics} might
% recognise a familiar pattern. We discuss this in~\S\ref{sec-related-optics}.}

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{text}
A1: 10     B1: A1 + A2
A2: 20     B2: B1 * 2
\end{minted}
\vspace{1mm}

\noindent
Here cell \cmd{A1} contains the value \cmd{10}, cell \cmd{B1} contains the
formula \cmd{A1 + A2}, etc. We can represent the formulae of this spreadsheet
with the following task description:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
sprsh1 :: Tasks Applicative String Integer
sprsh1 "B1" = Just $ Task $ \fetch -> ((+)  <$> fetch "A1"
                                            <*> fetch "A2")
sprsh1 "B2" = Just $ Task $ \fetch -> ((*2) <$> fetch "B1")
sprsh1 _    = Nothing
\end{minted}
\vspace{1mm}

\noindent
We instantiate keys \hs{k} with \hs{String}, and values \hs{v} with \hs{Integer}.
(Real spreadsheet cells would contain a wider range of values, of course.) The
task description \hs{sprsh1} embodies all the \emph{formulae} of the spreadsheet,
but not the input values. It pattern-matches on the key to see if it has a task
description (in the \Excel case, a formula) for it. If not, it returns
\hs{Nothing}, indicating that the key is an input. If there is a formula in the
cell, it returns the \hs{Task} to compute the value of the formula. Every
task is given a \emph{callback} \hs{fetch} to find the value of any keys on
which it depends. To run a \hs{Task}, we simply apply the function it holds to a
suitable callback (see the definition of the function \hs{run} in
Fig.~\ref{fig-types}).

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Applicative functors
pure  :: Applicative f => a -> f a
(<$>) :: Functor     f =>   (a -> b) -> f a -> f b -- Left-associative
(<*>) :: Applicative f => f (a -> b) -> f a -> f b -- Left-associative
\end{minted}
\vspace{0.5mm}
\begin{minted}[fontsize=\small]{haskell}
-- Standard State monad from Control.Monad.State
data State s a
instance Monad (State s)
get       :: State s s
gets      :: (s -> a) -> State s a
put       :: s -> State s ()
modify    :: (s -> s) -> State s ()
runState  :: State s a -> s -> (a, s)
execState :: State s a -> s -> s
\end{minted}
\vspace{0.5mm}
\begin{minted}[fontsize=\small]{haskell}
-- Standard types from Data.Functor.Identity and Data.Functor.Const
newtype Identity a = Identity { runIdentity :: a }
newtype Const m a  = Const    { getConst    :: m }
\end{minted}
\vspace{0.5mm}
\begin{minted}[fontsize=\small]{haskell}
instance Functor (Const m) where
    fmap _ (Const m) = Const m
\end{minted}
\vspace{0.5mm}
\begin{minted}[fontsize=\small]{haskell}
instance Monoid m => Applicative (Const m) where
    pure _              = Const mempty   -- mempty is identity for monoid m
    Const x <*> Const y = Const (x <> y) -- <> is the binary operation for m
\end{minted}
\vspace{0.5mm}
\begin{minted}[fontsize=\small]{haskell}
-- Standard types from Control.Monad.Trans.Writer
newtype WriterT w m a = WriterT { runWriterT :: m (a, w) }
tell :: Monad m => w   -> WriterT w m () -- write a value to the log
lift :: Monad m => m a -> WriterT w m a  -- lift an action into WriterT
\end{minted}
% \vspace{-3mm}
\caption{Standard library definitions.}\label{fig-stdlib}
% \vspace{-4mm}
\end{figure}

The code to ``compute the value of a formula'' in \hs{sprsh1} looks a bit
mysterious because it takes place in an \hs{Applicative}
computation~\cite{mcbride2008applicative} -- the relevant type signatures are
given in Fig.~\ref{fig-stdlib}. We will explain why in~\S\ref{sec-general-build}.
For now, we content ourselves with observing that a task description, of type
\hs{Tasks}~\hs{c}~\hs{k}~\hs{v}, is completely isolated from the world of
compilers, calc chains, file systems, caches, and all other complexities of real
build systems. It just computes a single output,
using a callback (\hs{fetch}) to find the values of its dependencies,
and limiting side effects to those described by \hs{c}.

\subsection{The Build Abstraction}\label{sec-general-build}

Next comes our second main abstraction -- a build system:

\vspace{0.5mm}
\begin{minted}[xleftmargin=10pt]{haskell}
type Build c i k v = Tasks c k v -> k -> Store i k v -> Store i k v
\end{minted}
\vspace{0.5mm}

\noindent
The signature is very straightforward. Given a task description, a target key,
and a store, the build system returns a new store in which the value of the
target key is up to date. What exactly does ``up to date'' mean?  We answer
that precisely in \S\ref{sec-build-correctness}.

\noindent
Here is a simple build system:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
busy :: Eq k => Build Applicative () k v
busy tasks key store = execState (fetch key) store
  where
    fetch :: k -> State (Store () k v) v
    fetch k = case tasks k of
        Nothing   -> gets (getValue k)
        Just task -> do v <- run task fetch
                        modify (putValue k v)
                        return v
\end{minted}
\vspace{1mm}

\noindent
The \hs{busy} build system defines the callback \hs{fetch} that, when given a
key, brings the key up to date in the store, and returns its value.
The function \hs{fetch} runs in the standard~Haskell \hs{State} monad
(Fig.~\ref{fig-stdlib}) initialised with the incoming \hs{store} by \hs{execState}.
To bring a key~\hs{k} up to date, \hs{fetch} asks the task description \hs{tasks}
how to compute its value. If \hs{tasks} returns \hs{Nothing} the key is an
input, so \hs{fetch} simply reads the result from the store. Otherwise \hs{fetch}
runs the obtained \hs{task} to produce a resulting value~\hs{v}, records the new
key/value mapping in the store, and returns \hs{v}. Notice that \hs{fetch}
passes itself to \hs{task} as an argument, so the latter can use \hs{fetch}
to recursively find the values of \hs{k}'s dependencies.

Given an acyclic task description, the \hs{busy} build system terminates with a
correct result, but it is not a \emph{minimal} build system
(Definition~\ref{def-minimal}). Since \hs{busy} has no memory
(\hs{i}~\hs{=}~\hs{()}), it cannot keep track of keys it has already built, and
will therefore busily recompute the same keys again and again if they have
multiple dependents. We will develop much more efficient build systems
in~\S\ref{sec-implementations}.

Nevertheless, \hs{busy} can easily handle the example task description
\hs{sprsh1} from the previous subsection~\S\ref{sec-task}. In the GHCi session
below we initialise the store with \cmd{A1} set to~10 and all other cells set
to~20.

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
@\ghci@ store  = initialise () (\key -> if key == "A1" then 10 else 20)
@\ghci@ result = busy sprsh1 "B2" store
@\ghci@ getValue "B1" result
30
@\ghci@ getValue "B2" result
60
\end{minted}
\vspace{1mm}

\noindent
As we can see, \hs{busy} built both \cmd{B2} and its dependency \cmd{B1} in the
correct order (if it had built \cmd{B2} before building \cmd{B1}, the result
would have been $20 * 2 = 40$ instead of $(10 + 20) * 2 = 60$). As an example
showing that \hs{busy} is not minimal, imagine that the formula in cell \cmd{B2}
was \cmd{B1~+~B1} instead of \cmd{B1~*~2}. This would lead to calling
\hs{fetch}~\hs{"B1"} twice -- once per occurrence of \cmd{B1} in the
formula -- and each call would recompute the formula in \cmd{B1}.

To avoid the recomputation, \hs{busy} can keep the set of processed keys in the
state monad (in addition to the \hs{store}), treat processed keys as inputs in
the \hs{getValue} branch of the \hs{fetch} callback, and include the key~\hs{k}
into the set of processed keys in the \hs{putValue} branch. This eliminates
unnecessary work \emph{within a single build}, but the next build needs to
recursively recompute all target's dependencies again even if no inputs changed.
To save work \emph{between builds}, it is necessary to store some build
information~\hs{i} persistently.

\subsection{The Need for Polymorphism in Task}\label{sec-why-polymorphism}

The previous example illustrates why the \hs{Task} abstraction is polymorphic
in~\hs{f}. Recall its definition from~\S\ref{sec-task}:

\begin{minted}[xleftmargin=10pt]{haskell}
newtype Task c k v = Task (@\std{forall}@ f. c f => (k -> f v) -> f v)
\end{minted}

\noindent
The \hs{busy} build system instantiates \hs{f} to
\hs{State}~\hs{(Store}~\hs{i}~\hs{k}~\hs{v)},
so that \hs{fetch}~\hs{::}~\hs{k}~\hs{->}~\hs{f}~\hs{v} can side-effect the
\hs{Store}, thereby allowing successive calls to \hs{fetch} to communicate with
one another.

We really, really want \hs{Task} to be \emph{polymorphic} in \hs{f}. Given
\emph{one} task description \cmd{T}, we want to explore \emph{many} build
systems that can build \cmd{T} -- and we will do so in
section~\S\ref{sec-implementations}. As we shall see, each build system will use
a different \hs{f}, so the task description must not fix~\hs{f}.

But the task description cannot possibly work for \emph{any} \hs{f} whatsoever; most task
descriptions (e.g. \hs{sprsh1} in \S\ref{sec-task}) require that \hs{f}
satisfies certain properties, such as \hs{Applicative} or \hs{Monad}. That is
why \hs{Task} has the ``\hs{c}~\hs{f}~\hs{=>}'' constraint in its type,
expressing that \hs{f} can only be instantiated by types that satisfy the
constraint \hs{c} and, in exchange, the task has access to the operations of class \hs{c}.
So the type \hs{Task} emerges naturally, almost inevitably.
But now that it \emph{has} emerged, we find that constraints \hs{c} classify
task descriptions in a very interesting, and practically useful, way:

\begin{itemize}
\item \hs{Task}~\hs{Applicative}: In \hs{sprsh1} we needed only \hs{Applicative}
  operations, expressing the fact that the dependencies between cells can be
  determined \emph{statically}; that is, by looking at the formulae, without
  ``computing'' them (see \S\ref{sec-deps}).
\item \hs{Task}~\hs{Monad}: As we shall see in \S\ref{sec-task-monad}, a monadic
  task allows \emph{dynamic} dependencies, in which a formula may depend on cell
  \cmd{C}, but \emph{which} cell \cmd{C} depends on the value of another cell
  \cmd{D}. A simple example of a task with dynamic dependencies is \Excel
  formula \cmd{INDIRECT("A" \& C1)} from~\S\ref{sec-background-excel}.
\item \hs{Task}~\hs{Functor} is somewhat degenerate: a functorial task
  description cannot even use the application operator \hs{<*>}, which limits
  dependencies to a linear chain, as e.g. in Docker containers~\cite{docker}
  (ignoring the recent multi-stage builds). It
  is interesting to note that, when run on such a task description, the
  \hs{busy} build system will build each key at most once, thus partially
  fulfilling the minimality requirement~\ref{def-minimal}. Alas, it still has no
  mechanism to decide which input keys changed since the previous build.
\item \hs{Task}~\hs{Selective} corresponds to task descriptions with
  \emph{conditional statements}, e.g. \Excel formula \cmd{IF(C1=1,B2,A2)},
  where it is possible to statically \emph{over-approximate} the set of task
  dependencies. Here \hs{Selective} is a type class of \emph{selective
  applicative functors}~\cite{mokhov_selective_2019}, which allows us to model
  build systems like \Dune~\cite{dune} using the presented framework.
\item \hs{Task}~\hs{MonadFail} corresponds to monadic tasks that may fail. For
  example, the formula \cmd{A1/A2} may fail due to division by zero. We will
  discuss this in~\S\ref{sec-failures}.
\item \hs{Task}~\hs{MonadPlus}, \hs{Task}~\hs{MonadRandom} and their
  variants can be used for describing tasks with a certain type of
  non-determinism, as discussed in~\S\ref{sec-non-determinism}.
\item \hs{Task}~\hs{(MonadState}~\hs{i)} will be used
  in~\S\ref{sec-implementations} to describe tasks that have read and write
  access to the persistently stored build information~\hs{i}.
\end{itemize}

% \vspace{-0.5mm}
\subsection{Monadic Tasks}\label{sec-task-monad}

As explained in~\S\ref{sec-background-excel}, some task descriptions have
dynamic dependencies, which are determined by values of intermediate
computations. In our framework, such task descriptions correspond to the type
\hs{Task}~\hs{Monad}~\hs{k}~\hs{v}. Consider this spreadsheet example:

\vspace{0.5mm}
\begin{minted}[xleftmargin=10pt]{text}
A1: 10      B1: IF(C1=1,B2,A2)      C1: 1
A2: 20      B2: IF(C1=1,A1,B1)
\end{minted}
\vspace{0.5mm}

\noindent
Note that \cmd{B1} and \cmd{B2} statically form a dependency cycle, so \Make
would not be able to order the tasks topologically, but \Excel, which uses
dynamic dependencies, is perfectly happy. We can express this spreadsheet using
our task abstraction as follows:

% The spreadsheet example that uses
% the \hs{INDIRECT} function can be expressed very similarly: simply replace the
% line containing the \cmd{if} statement with \hs{fetch ("A" ++ show c1)}.

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
sprsh2 :: Tasks Monad String Integer
sprsh2 "B1" = Just $ Task $ \fetch -> do
    c1 <- fetch "C1"
    if c1 == 1 then fetch "B2" else fetch "A2"
sprsh2 "B2" = Just $ Task $ \fetch -> do
    c1 <- fetch "C1"
    if c1 == 1 then fetch "A1" else fetch "B1"
sprsh2 _ = Nothing
\end{minted}
\vspace{1mm}

\noindent
The big difference compared to \hs{sprsh1} is that the computation now takes
place in a \hs{Monad}, which allows us to extract the value of \hs{c1} and
\hs{fetch} \emph{different keys} depending on whether or not
\hs{c1}~\hs{==}~\hs{1}. Note that in this example one can statically determine
the sets of possible dependencies of the formulae \cmd{IF(C1=1,B2,A2)} and
\cmd{IF(C1=1,A1,B1)} but this cannot be done in general -- recall the
spreadsheet with the formula \cmd{INDIRECT("A" \& C1)}
from~\S\ref{sec-background-excel}, where the argument of the \cmd{INDIRECT}
function is a string computed dynamically during the build. Such tasks can also
be captured using \hs{Tasks}~\hs{Monad}:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
sprsh3 :: Tasks Monad String Integer
sprsh3 "B1" = Just $ Task $ \fetch -> do
    c1 <- fetch "C1"
    fetch ("A" ++ show c1)
sprsh3 _ = Nothing
\end{minted}
\vspace{1mm}

\noindent
Since the \hs{busy} build system introduced in~\S\ref{sec-general-build} always
rebuilds every dependency it encounters, it is easy for it to handle dynamic
dependencies. For minimal build systems, however, dynamic dependencies, and hence
monadic tasks, are much more challenging, as we shall see
in~\S\ref{sec-implementations}.

\subsection{Correctness of a Build System}\label{sec-build-correctness}

We can now say what it means for a build system to be \emph{correct}, something
that is seldom stated formally. Our intuition is this: \emph{when the build
system completes, the target key, and all its dependencies, should be up to
date}. What does ``up to date'' mean? It means that if we recompute the value of
the key (using the task description, and the final store), we should get exactly
the same value as we see in the final store.

To express this formally we need an auxiliary function \hs{compute}, that
computes the value of a key in a given store \emph{without attempting to update
any dependencies}:

\begin{minted}[xleftmargin=10pt]{haskell}
compute :: Task Monad k v -> Store i k v -> v
compute task store = runIdentity (run task fetch)
  where
    fetch :: k -> Identity v
    fetch k = Identity (getValue k store)
\end{minted}

\noindent
Here we do not need any effects in the \hs{fetch} callback to \hs{task}, so
we can use the standard Haskell \hs{Identity} monad (Fig.~\ref{fig-stdlib}).
This is another use of polymorphism in \hs{f}, discussed in~\S\ref{sec-why-polymorphism}.
The use of \hs{Identity} just fixes the ``impedance mismatch'' between the
function \hs{getValue}, which returns a pure value~\hs{v}, and the \hs{fetch}
argument of the \hs{task}, which must return an \hs{f}~\hs{v} for some \hs{f}.
To fix the mismatch, we wrap the result of \hs{getValue} in the \hs{Identity}
monad and pass to the \hs{task}. The result has type \hs{Identity}~\hs{v}, which
we unwrap with \hs{runIdentity}.

\vspace{2mm}
\textbf{Definition}~(Correctness). Suppose \hs{build} is a build system,
\hs{tasks} is a build task description, \hs{key} is a target key, \hs{store} is
an initial store, and \hs{result} is the store produced by running the build
system with parameters \hs{tasks}, \hs{key} and \hs{store}. Or, using the
precise language of our abstractions:

\begin{minted}[xleftmargin=10pt]{haskell}
build         :: Build c i k v
tasks         :: Tasks c k v
key           :: k
store, @@result :: Store i k v
result = @@build @@tasks @@key @@store
\end{minted}

\noindent
The keys that are reachable from the target \hs{key} via dependencies fall
into two classes: input keys and non-input keys, which we will denote by $I$ and
$O$, respectively. Note that \hs{key} may be in either of these sets, although
the case when \hs{key} is an input is degenerate: we have $I=\{$\hs{key}$\}$ and
$O=\emptyset$.

The build \hs{result} is~\emph{correct} if the following two conditions hold:

\begin{itemize}
    \item \hs{result} and \hs{store} \emph{agree on inputs}, that is, for all
    input keys \hs{k}~$\in$~$I$:
    \[
    \hs{getValue}~\hs{k}~\hs{result}~\hs{==}~\hs{getValue}~\hs{k}~\hs{store}.
    \]
    In other words, no inputs were corrupted during the build.

    \item The \hs{result} is \emph{consistent} with the \hs{tasks}, i.e.
    for all non-input keys \hs{k}~$\in$~$O$, the result of recomputing the
    corresponding \hs{task} matches the value stored in the \hs{result}:
    \[
    \hs{getValue}~\hs{k}~\hs{result}~\hs{==}~\hs{compute}~\hs{task}~\hs{result}.
    \]
\end{itemize}
A build system is \emph{correct} if it produces a correct \hs{result} for any
\hs{tasks}, \hs{key} and \hs{store}.

\label{def-correct}
\vspace{2mm}

It is hard to satisfy the above definition of correctness given a task
description with cycles. All build systems discussed in this paper are correct
only under the assumption that the given task description is acyclic. This
includes the \hs{busy} build system introduced earlier: it will loop
indefinitely given a cyclic \hs{tasks}. Some build systems provide a limited
support for cyclic tasks, see~\S\ref{sec-iterative-compute}.

The presented definition of correctness needs to be adjusted for build systems
that support non-deterministic tasks and shallow cloud builds, as will be
discussed in sections~\S\ref{sec-non-determinism} and~\S\ref{sec-cloud-aspects},
respectively.

\subsection{Computing Dependencies}\label{sec-deps}

Earlier we remarked that a \hs{Task}~\hs{Applicative} could only have static
dependencies. Usually we would extract such static dependencies by (in the case
of \Excel) looking at the syntax tree of the formula.  But a task description
has no such syntax tree: as you can see in the definition of \hs{Task} in
Fig.~\ref{fig-types}, a task is just a function, so all we can do is call it.
Yet, remarkably, we can use the polymorphism of a
\hs{Task}~\hs{Applicative} to find its dependencies \emph{without doing any of
the actual work}. Here is the code:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
dependencies :: Task Applicative k v -> [k]
dependencies task = getConst $ run task (\k -> Const [k])
\end{minted}
\vspace{1mm}

\noindent
Here \hs{Const} is a standard Haskell type defined in Fig.~\ref{fig-stdlib}. We
instantiate \hs{f} to \hs{Const}~\hs{[@@k]}. So a value of type \hs{f}~\hs{v},
or in this case \hs{Const}~\hs{[@@k]}~\hs{v}, contains no value \hs{v}, but does
contain a list of keys of type \hs{[@@k]} which we use to record dependencies.
The \hs{fetch} callback that we pass to \hs{task} records a single dependency;
and the standard definition of \hs{Applicative} for \hs{Const} (which we give
in Fig.~\ref{fig-stdlib}) combines the dependencies from different parts of the
task. Running the task with \hs{f}~=~\hs{Const}~\hs{[@@k]} will thus
accumulate a list of the task's dependencies -- and that is what
\hs{dependencies} does:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
@\ghci@ dependencies $ fromJust $ sprsh1 "B1"
["A1","A2"]
\end{minted}
\begin{minted}[xleftmargin=10pt]{haskell}
@\ghci@ dependencies $ fromJust $ sprsh1 "B2"
["B1"]
\end{minted}
\vspace{1mm}

\noindent
Notice that these calls to \hs{dependencies} do no actual computation (in this
case, spreadsheet arithmetic). They cannot: we are not supplying a store or any
input numbers. So, through the wonders of polymorphism, we are able to extract
the dependencies of the spreadsheet formula, and to do so efficiently, simply by
running its code in a different \hs{Applicative}! This is not new, for example
see~Capriotti~\&~Kaposi~\shortcite{free-applicatives}, but it is extremely cool.
We will see a practical use for \hs{dependencies} when implementing applicative
build systems, see~\S\ref{sec-implementation-make}.

So much for applicative tasks. What about monadic tasks with dynamic
dependencies? As we have seen in~\S\ref{sec-background-shake}, dynamic
dependencies need to be tracked too. This cannot be done statically; notice that
we cannot apply the function \hs{dependencies} to a \hs{Task}~\hs{Monad} because
the \hs{Const} functor has no \hs{Monad} instance. We need to run a monadic task
on a store with concrete values, which will determine the discovered
dependencies. Accordingly, we introduce the function \hs{track} -- a combination
of \hs{compute} and \hs{dependencies} that computes both the resulting value and
the list of its dependencies (key/value pairs) in an arbitrary monadic
context~\hs{m}. We need this function to be polymorphic over~\hs{m}, because
each build system will execute tasks in its own monad, as we shall see
in~\S\ref{sec-implementations}.

Here is an implementation of \hs{track} based on the standard Haskell
\hs{WriterT} \emph{monad transformer}~\cite{liang1995monad}, whose main
types are listed in Fig.~\ref{fig-stdlib}:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
track :: Monad m => Task Monad k v -> (k -> m v) -> m (v, [(k, v)])
track task fetch = runWriterT $ run task trackingFetch
  where
    trackingFetch :: k -> WriterT [(k, v)] m v
    trackingFetch k = do v <- lift@\,@(fetch k); tell@\,@[(k, v)]; return@\,\blk{v}@
\end{minted}
\vspace{1mm}

\noindent
This function uses the \hs{WriterT} transformer for recording additional
information~--~a list of key/value pairs \hs{[(@@k,}~\hs{v)]}~--~when executing
a task in an arbitrary monad~\hs{m}. We substitute the given \hs{fetch} with a
\hs{trackingFetch} that, in addition to fetching a value, tracks the
corresponding key/value pair. The \hs{task} returns a value of type
\hs{WriterT}~\hs{[(@@k,}~\hs{v)]}~\hs{m}~\hs{v}, which we unwrap with
\hs{runWriterT}. We will use \hs{track} when implementing monadic build systems
with dynamic dependencies, see~\S\ref{sec-implementation-shake}.

Here we show an example of \hs{track}ing monadic tasks when
\hs{m}~\hs{=}~\hs{IO}, by defining a corresponding \hs{fetchIO} of type
\hs{String}~\hs{->}~\hs{IO}~\hs{Integer}, which allows us to demonstrate the
dynamic nature of monadic dependencies in GHCi.

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
@\ghci@ fetchIO k = do putStr (k ++ ": "); read <$> getLine
@\ghci@ track (fromJust $ sprsh2 "B1") fetchIO
C1: 1
B2: 10
(10,[("C1",1),("B2",10)])
\end{minted}
\vspace{-2mm}
\begin{minted}[xleftmargin=10pt]{haskell}
@\ghci@ track (fromJust $ sprsh2 "B1") fetchIO
C1: 2
A2: 20
(20,[("C1",2),("A2",20)])
\end{minted}
\vspace{1mm}

\noindent
As expected, the dependencies of the cell \cmd{B1} from \hs{sprsh2} (see the
spreadsheet in~\S\ref{sec-task-monad}) are determined by the value of \cmd{C1},
which in this case is obtained by reading from the standard input using
\hs{fetchIO}.

\subsection{Examples of Tasks}\label{sec-task-examples}

In this section we give examples of tasks whose definitions involve different
constraints on the computation context: \hs{Functor}, \hs{Applicative},
\hs{Monad} and \hs{MonadState}~\hs{s}. The purpose of these examples is to
continue building the intuition behind the \hs{Task} abstraction, and prepare
the reader for richer types of tasks that will appear
in~\S\ref{sec-implementations} and~\S\ref{sec-engineering}.

We start with one of the favourite examples for functional
programmers -- the \emph{Fibonacci sequence} $F_n = F_{n-1} + F_{n-2}$:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
fibonacci :: Tasks Applicative Integer Integer
fibonacci n = if n < 2 then Nothing else
    Just $ Task $ \fetch -> (+) <$> fetch (n - 1) <*> fetch (n - 2)
\end{minted}
\vspace{1mm}

Here the keys $n < 2$ are input parameters, and one can obtain the usual
Fibonacci sequence by picking $F_0=0$ and $F_1=1$, respectively. Any minimal
build system will compute the sequence with memoization, i.e. without
recomputing the same value twice.

% TODO: Add a figure with a dependency graph?
Dependencies of elements of the Fibonacci sequence are known statically, hence
we can express it using \hs{Tasks}~\hs{Applicative}, and benefit from static
dependency analysis (\S\ref{sec-deps}):

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
@\ghci@ dependencies (fromJust $ fibonacci 5)
[4,3]
\end{minted}
\vspace{1mm}

\noindent
Interestingly, the \emph{Ackermann function} -- a famous example of a function
that is not primitive recursive -- cannot be expressed as an applicative task,
because it needs to perform an intermediate recursive call to determine the
value of one of its dependencies $A(m, n-1)$:

\[
A(m, n) = \left\{
\begin{array}{ll}
n+1 & \mbox{if } m = 0 \\
A(m-1, 1) & \mbox{if } m > 0 \mbox{ and } n = 0 \\
A(m-1, A(m, n-1)) & \mbox{if } m > 0 \mbox{ and } n > 0.
\end{array}\right.
\]

We therefore use \hs{Tasks}~\hs{Monad} to express this function, binding the
dynamic dependency to variable \hs{index}:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
ackermann :: Tasks Monad (Integer, Integer) Integer
ackermann (m, n)
  | m < 0 || n < 0 = Nothing
  | m == 0    = Just $ Task $ const $ pure (n+1)
  | n == 0    = Just $ Task $ \fetch -> fetch (m-1, 1)
  | otherwise = Just $ Task $ \fetch -> do index <- fetch (m, n-1)
                                           fetch (m-1, index)
\end{minted}
\vspace{1mm}

\noindent
Functorial tasks are less common than applicative and monadic, but there is a
classic example too -- the \emph{Collatz sequence}, where given an initial value
$c_0$, we calculate the next value $c_n$ from $c_{n-1}$ either by dividing
$c_{n-1}$ by 2 (if it is even) or multiplying it by 3 and adding 1 (if it is
odd):

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
collatz :: Tasks Functor Integer Integer
collatz n | n <= 0    = Nothing
          | otherwise = Just $ Task $ \fetch -> f <$> fetch (n - 1)
  where
    f k | even k    = k `div` 2
        | otherwise = 3 * k + 1
\end{minted}
\vspace{1mm}

\noindent
Functorial tasks correspond to computations with a linear dependency chain. For
example, computing the element $c_8$ of the Collatz sequence starting from
$c_0=6$ leads to the following dependency chain:
$c_0 = 6 \rightarrow 3 \rightarrow 10 \rightarrow 5 \rightarrow 16 \rightarrow 8
\rightarrow 4 \rightarrow 2 \rightarrow 1 = c_8$.

Collatz sequence is a good example of the early cutoff optimisation
(\S\ref{sec-background-shake}): if we recompute $c_8$ starting from a different
initial value $c_0=40$, the resulting computation will have a large overlap with
the previous one:
$c_0 = 40 \rightarrow 20 \rightarrow 10 \rightarrow 5 \rightarrow 16 \rightarrow
8 \rightarrow 4 \rightarrow 2 \rightarrow 1 = c_8$. We can therefore stop the
recomputation after just two steps, since $c_2=10$ has not changed.

Note that we can statically extract even more precise dependency information
from functorial tasks compared to applicative tasks. Indeed, we statically know
that a \hs{Task}~\hs{Functor} has \emph{exactly one} \hs{dependency}:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
dependency :: Task Functor k v -> k
dependency task = getConst (run task Const)
\end{minted}
\vspace{1mm}

\noindent
The \hs{Tasks} abstraction allows us to express pure functions in a way that
is convenient for their memoization and incremental recomputation
(see~\S\ref{sec-related-memo} for a discussion on memoization). If we
furthermore need to share computation results via a cloud cache, we can use
\hs{Tasks}~\hs{(MonadState}~\hs{s)} that will
play an important role in~\S\ref{sec-implementations}. Intuitively, by making
a shared state of type \hs{s} available to a task, we give it the abilities to
lookup and update cached computation results using the \hs{MonadState} methods
\hs{get} and \hs{modify}. For example, below we implement a cloud version of the
Ackermann task that uses a \hs{Cache} of type
\hs{Map}~\hs{(Integer,}~\hs{Integer)}~\hs{Integer} for sharing results of known
Ackermann values.

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
type Cache = Map (Integer, Integer) Integer

cloudAckermann :: Tasks (MonadState@\,@Cache)@\,@(Integer,@\,@Integer)@\,@Integer
cloudAckermann (m, n)
    | m < 0 || n < 0 = Nothing
    | m == 0    = Just $ Task $ const $ pure (n+1)
    | n == 0    = Just $ Task $ \fetch -> fetch (m-1, 1)
    | otherwise = Just $ Task $ \fetch -> do
        cache <- get
        case Map.lookup (m, n) cache of
            Nothing -> do index <- fetch (m, n-1)
                          value <- fetch (m-1, index)
                          modify (Map.insert (m, n) value)
                          return value
            Just value -> return value
\end{minted}
\vspace{1mm}

\noindent
The main case ($m>0 \wedge n>0$) starts by looking up the pair of indices
\hs{(}\hs{m,}~\hs{n)} in the \hs{cache}. If the cache has \hs{Nothing}, we
calculate the resulting \hs{value} as before and \hs{modify} the cache
accordingly; otherwise, if we have a cache hit, we return the obtained value
immediately, skipping the actual calculation and thus potentially saving a large
amount of work. Indeed you do not want to recompute $A(4,2)=2^{65536}-3$
unnecessarily; all of its 19,729 decimal digits have already been helpfully
computed, e.g. see~Kosara~\shortcite{ackermann42}. We will use \hs{MonadState}
tasks in our models of cloud build systems in~\S\ref{sec-implementations}.