### 2019-10-20 papers/jfp/8-engineering.tex
Final tweaks

\section{Engineering Aspects}\label{sec-engineering}

In the previous sections we have modelled the most critical subset of various
build systems. However, like all real-world systems, there are many corners that
obscure the essence. In this section we discuss some of those details, what
would need to be done to capture them in our model, and what the impact would be.

\subsection{Partial Stores and Exceptions}\label{sec-failures}

Our model assumes a world where the store is fully-defined, every \hs{k} is
associated with a \hs{v}, and every compute successfully completes returning a
valid value. In the real world, build systems frequently deal with errors, e.g.
``file not found'' or ``compilation failed''. There are many ways of modelling
errors, and in this section we give three simple examples.

% We can model such failure
% conditions by instantiating \hs{v} to either \hs{Maybe}~\hs{v} (for missing
% values) or \hs{Either}~\hs{e}~\hs{v} (for exceptions of type \hs{e}). That can
% model the errors inside the store and the \hs{Task}, but because our build
% system models are polymorphic in \hs{v}, they cannot apply any special behaviour
% for errors, e.g. aborting the build early.

One simple approach is to include failures into the type of values \hs{v}, for
example, to model a partial store we can use an algebraic data type isomorphic
to \hs{Maybe}:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
data Value = FileNotFound | FileContents String
\end{minted}
\vspace{1mm}

\noindent
This is convenient if \emph{tasks are aware of failures}. For example, a task
may be able to cope with missing files, e.g. if \hs{fetch}~\hs{"username.txt"}
returns \hs{FileNotFound}, the task could use the literal string \hs{"User"} as
a default value. In this case the task will \emph{depend} on the fact that the
file \cmd{username.txt} is missing, and will need to be rebuilt if the user
later creates this file. In general, we can use values of type
\hs{Either}~\hs{e}~\hs{v} when dealing with failures of type \hs{e}. To
automatically convert a ``failure-free'' task into an equivalent task operating
with values of type \hs{Either}~\hs{e}~\hs{v} we can use the following function:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
liftEither :: Task Monad k v -> Task Monad k (Either e v)
liftEither task = Task $ \fetch ->
    runExceptT $ run task (ExceptT . fetch)
\end{minted}
\vspace{1mm}

\noindent
Here \hs{liftEither} wraps the result of the given
\hs{fetch}~\hs{::}~\hs{k}~\hs{->}~\hs{Either}~\hs{e}~\hs{v} into \hs{ExceptT}
and then runs the task in the \hs{ExceptT} monad
transformer~\cite{liang1995monad}.

Another approach is to include failures into the computation context \hs{f}.
Recall from~\S\ref{sec-why-polymorphism} that we require tasks to be polymorphic
in~\hs{f}, and can therefore choose to execute a \hs{Task} in an~\hs{f} with
failures, the simplest example being \hs{f}~\hs{=}~\hs{Maybe}. Below we define
a callback that returns \hs{Just}~\hs{value} for keys \cmd{A1} and \cmd{A2} but
fails with \hs{Nothing} on all other keys:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
fetchA1A2 :: String -> Maybe Integer
fetchA1A2 k = Map.lookup k (Map.fromList [("A1", 10), ("A2", 20)])
\end{minted}
\vspace{1mm}

\noindent
We can directly run any \hs{Task} with this callback. For example, the task
\cmd{B1 = A1 + A2} from \hs{sprsh1} in~\S\ref{sec-task} returns
\hs{Just}~\hs{30}, whereas the task \cmd{B2 = B1 * 2} returns \hs{Nothing}
because \hs{fetchA1A2} fails on \cmd{B1}.

This approach is convenient if \emph{tasks are not aware of failures}, e.g. we
can model \Excel formulae as pure arithmetic functions, and introduce failures
``for free'' if/when needed by instantiating \hs{Tasks} with an appropriate
\hs{f}. In a real system this \hs{f} would be more complex than just \hs{Maybe},
for example \hs{MaybeT}~\hs{(State}~\hs{Spreadsheet)}, thus allowing us to
combine failures with access to \Excel's spreadsheet state by using the
\hs{MaybeT} monad transformer~\cite{liang1995monad}.

Finally, the task itself might not want to encode failures into the type of
values \hs{v}, but instead \emph{demand that} \hs{f} \emph{has a built-in notion
of failures}. This can be done by choosing a suitable constraint \hs{c}, such as
\hs{Alternative}, \hs{MonadPlus} or even better something specific to failures,
such as \hs{MonadFail}. Then both the callback and the task can reuse the same
failure mechanism as shown below:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
class Monad m => MonadFail m where
    fail :: String -> m a
\end{minted}
\begin{minted}[xleftmargin=10pt]{haskell}
sprsh4 :: Tasks MonadFail String Integer
sprsh4 "B1" = Just $ Task $ \fetch -> do
    a1 <- fetch "A1"
    a2 <- fetch "A2"
    if a2 == 0 then fail "division by 0" else return (a1 `div` a2)
sprsh4 _ = Nothing
\end{minted}
\vspace{1mm}

\noindent
With this approach we can implement a build system that accepts
\hs{Tasks}~\hs{MonadFail}~\hs{k}~\hs{v} and handles errors by aborting the
build early and returning
\hs{Either}~\hs{String}~\hs{(Store}~\hs{i}~\hs{k}~\hs{v)} instead of just
\hs{Store}~\hs{i}~\hs{k}~\hs{v} as in our \hs{Build}
abstraction~\S\ref{sec-general-build}. One possible implementation of such a
failure-handling build system is based on adding an extra
\hs{Either}~\hs{String} layer into the monad stack, e.g. augmenting the monad
\hs{State}~\hs{(Store}~\hs{i}~\hs{k}~\hs{v)} used by the schedulers
in~\S\ref{sec-implementations} with exceptions. We omit the actual
implementation: while fairly direct it is also tedious due to additional
wrapping and unwrapping.

% \todo{AM}{Add a MonadFail build system.}

\subsection{Parallelism}\label{sec-parallelism}

We have given simple implementations assuming a single thread of execution,
but all the build systems we address can actually build independent keys in
parallel. While it complicates the model, the complications are restricted
to the scheduler:

\begin{enumerate}
\item The \hs{topological} scheduler can build the full dependency graph, and
whenever all dependencies of a task are complete, the task itself can be
started.

\item The \hs{restarting} scheduler can be made parallel in a few ways, but the
most direct is to have $n$ threads reading keys from the build queue. As before,
if a key requires a dependency not yet built, it is moved to the end~--~the
difference is that sometimes keys will be moved to the back of the queue not
because they are out of date but because of races with earlier tasks that had
not yet finished. As a consequence, if the build order is persisted over
successive runs (as in \Excel), potentially racey dependencies will be
separated, giving better parallelism over time.

\item The \hs{suspending} scheduler can be made parallel by starting multiple
dependencies in parallel. One approach is to make the request for dependencies
take a list of keys, as implemented by \Shake. Another approach is to treat the
\hs{Applicative} dependencies of a \hs{Task}~\hs{Monad} in parallel, as
described by~Marlow~\etal~\shortcite{marlow2014haxl}.
\end{enumerate}

Once sufficient parallelism is available the next challenge is preventing excess
parallelism and machine resource starvation, which is usually achieved with a
thread pool/limit.

The actual implementation of the parallel schedulers is not overly onerous, but
neither is it beautiful or informative, so we omit it.

\subsection{Impure Computations}\label{sec-non-determinism}

In our model we define \hs{Task} as a function -- when given the same inputs
it will always produce the same output. Alas, the real-world is not so obliging.
Some examples of impure tasks include:

\begin{itemize}
\item \textbf{Untracked dependencies}: Some tasks depend on untracked values --
      for example C compilation will explicitly list the \cmd{source.c} file as
      a dependency, but it may not record that the version of \cmd{gcc} is also
      a dependency.

\item \textbf{Non-determinism}: Some tasks are \emph{non-deterministic},
      producing any result from a possible set. As an example, GHC programs
      compiled using parallelism can change the order in which unique variables
      are obtained from the name supply, producing different but semantically
      identical results.

\item \textbf{Volatility}: Some tasks are defined to change in every build, e.g.
      \Excel provides a~``function'' \cmd{RANDBETWEEN} producing a fresh random
      number in a specified range on each recalculation. Similarly, build
      systems like \Make and \Shake provide volatile \emph{phony rules}. The
      main difference from non-deterministic tasks is that volatile tasks cannot
      be cached. They are readily modelled as depending on a special key
      \cmd{RealWorld} whose value is changed in every build.
\end{itemize}

There is significant interplay between all three sources of
impurity. Systems like \Bazel use various sandboxing techniques to guard against
missing dependencies, but none are likely to capture all dependencies right down
to the CPU model and microcode version. Tasks with untracked dependencies can be
marked volatile, a technique \Excel takes with the \cmd{INDIRECT} function,
fixing the untracked dependency at the cost of minimality.

Most of the implementations in~\S\ref{sec-implementations} can deal with
non-determinism, apart from \Buck, which uses the assumption of task determinism
to reduce the number of cloud lookups: if all tasks are deterministic, one needs
just a single cloud lookup for obtaining the value of the target key from the
hash of its terminal inputs. This is highly desirable not only for shallow
builds, but for non-shallow builds too, because under the assumption of task
determinism \emph{all} intermediate values are fully determined by the terminal
inputs, and can therefore be requested in a single batch query to the server.

One way of modelling non-determinism is to \emph{demand that}~\hs{f}~\emph{has a
built-in source of non-determinism}, for example, by enriching
\hs{Tasks}~\hs{Monad} to \hs{Tasks}~\hs{MonadRandom} that has access to an
additional method
\hs{getRandom}~\hs{::}~\hs{(Integer,}~\hs{Integer)}~\hs{->}~\hs{m}~\hs{Integer},
i.e. a source of random numbers in a specified range. Here is a task description
corresponding to a spreadsheet with the formula \cmd{B1 = A1 + RANDBETWEEN(1,2)}:

\begin{minted}[xleftmargin=10pt]{haskell}
sprsh5 :: Tasks MonadRandom String Integer
sprsh5 "B1" = Just $ Task $ \fetch -> do a1 <- fetch "A1"
                                         r  <- getRandom (1,2)
                                         return (a1 + r)
sprsh5 _    = Nothing
\end{minted}
\vspace{1mm}

\noindent
Such tasks can be modelled in our framework by adjusting the correctness
definition~(\S\ref{sec-build-correctness}): instead of requiring that the
produced value \emph{equals the result} of recomputing the task, we now require
that produced value \emph{belongs to the set of possible results} of
recomputing the task, i.e. the set \cmd{$\{$A1 + 1, A1 + 2$\}$} in the above
example.

% \todo{AM}{Show implementation of \hs{computeND}.
% Neil says: not convinced it's that interesting.}.

Interestingly, \hs{Task}~\hs{MonadRandom} is powerful enough to express
dependency-level non-determinism, for example,
\cmd{INDIRECT("A" \& RANDBETWEEN(1,2))}, whereas most build tasks in real-life
build systems only experience a value-level non-determinism. \mbox{\Excel}
handles this example simply by marking the cell volatile~--~an approach that can
be readily adopted by any of our implementations.

\vspace{-1mm}
\subsection{Cloud Implementation}\label{sec-cloud-aspects}

Our model of cloud builds provides a basic framework to discuss and reason
about them, but lacks a number of important engineering corners:

\begin{itemize}
\item \textbf{Communication}: When traces or contents are stored in the cloud,
communication can become a bottleneck, so it is important to send only the
minimum amount of information, optimising with respect to build system
invariants. For example, incremental data processing systems in the cloud, such
as \Reflow~\cite{reflow}, need to efficiently orchestrate terabytes of data.

\item \textbf{Offloading}: Once the cloud is storing build products and traces,
it is possible for the cloud to also contain dedicated workers that can execute
tasks remotely~--~offloading some of the computation and potentially running
vastly more commands in parallel.

\item \textbf{Eviction}: The cloud storage, as modelled
in~\S\ref{sec-constructive-traces}, grows indefinitely, but often resource
constraints require evicting old items from the store. When evicting an old
value \hs{v}, one can also evict all traces mentioning the now-defunct
\hs{hash}~\hs{v}. However, for shallow builds (see below), it is beneficial to
keep these traces, allowing builds to ``pass-through'' hashes whose underlying
values are not known, recreating them only when they must be materialised.

\item \textbf{Shallow builds}: Building the end target, e.g. an installer
package, often involves many intermediate tasks. The values produced by these
intermediate tasks may be large, so some cloud build systems are designed to
build end targets \emph{without materialising} any intermediate values,
producing a so-called \emph{shallow build} -- see an example
in~\S\ref{sec-background-bazel}. Some build systems go even further, integrating
with the file system to only materialise the file when the user accesses
it~\cite{gvfs}.
\end{itemize}

Shallow builds have a slightly weaker correctness property than in the
Definition~\ref{def-correct}. Instead of demanding that \emph{all} keys reachable
from the target match, we only require matches for the target itself and the
\emph{input keys} reachable from the target.

\begin{figure}
\begin{subfigure}[b]{0.31\linewidth}
\centerline{\includegraphics[scale=0.27]{fig/frankenbuild-example-build.pdf}}
% \vspace{-1.5mm}
\caption{Initial build}
\end{subfigure}
\begin{subfigure}[b]{0.31\linewidth}
\centerline{\includegraphics[scale=0.27]{fig/frankenbuild-example-clean.pdf}}
% \vspace{-1.5mm}
\caption{Clean up, evict \cmd{main.prof}}
\end{subfigure}
\begin{subfigure}[b]{0.36\linewidth}
\centerline{\includegraphics[scale=0.27]{fig/frankenbuild-example-rebuild.pdf}}
% \vspace{-1.5mm}
\caption{Build \cmd{main.prof} and \cmd{report.txt}}
\end{subfigure}
\vspace{-1mm}
\caption{Frankenbuild example: (a)~build a human-readable profiling report for
\cmd{main.exe} from information dump \cmd{main.prof} produced by a profiling
tool, saving deep constructive traces in the cloud, (b)~remove built files
locally and evict \cmd{main.prof} from the cloud storage, (c)~rebuild
\cmd{main.prof} (profiling is non-deterministic, hence a new hash value), then
build \cmd{report.txt} by downloading it from the matching deep constructive
trace in the cloud. The result is a Frankenbuild because \cmd{main.prof} and
\cmd{report.txt} are inconsistent. New and evicted cloud storage entries are
highlighted; file hashes are shown in circles.
\label{fig-frankenbuild}}
% \vspace{-4mm}
\end{figure}

As described in \S\ref{sec-using-cloud-shake}, non-determinism
(\S\ref{sec-non-determinism}) is harmful for cloud builds, reducing the number
of cache hits. However, for deep constructive traces
(\S\ref{sec-deep-constructive-traces}) it is much worse, even leading to
\emph{incorrect results}. Fig.~\ref{fig-frankenbuild} shows a
\emph{Frankenbuild}~\cite{esfahani2016cloudbuild} example, where the target
\cmd{report.txt}, which is downloaded from the cloud, is inconsistent with its
immediate dependency \cmd{main.prof}. This inconsistency is caused by two factors:
(i)~inherent non-determinism of profiling -- running a profiling tool on the very
same \cmd{main.exe} will produce different \cmd{main.prof} results every time;
and (ii)~relying on deep constructive traces that cache build results based
on the hashes of terminal task inputs (in this case \cmd{main.exe}). The result
violates all three definitions of correctness: the main
definition~(\S\ref{sec-build-correctness}), the variant for non-deterministic
tasks~(\S\ref{sec-non-determinism}) and the variant for shallow
builds~(this section).

% \todo{AM/NM}{Re-using existing infrastructure: (a) key-value store, (b) remote
% execution service. Neil says: I think this is a stretch. It's a lot more nuanced
% as we saw from Bazel discussions so I'd probably skip it.}

\vspace{-1mm}
\subsection{Iterative Computations}\label{sec-iterative-compute}

Some computations are best described not by a chain of acyclic dependencies,
but by a loop. For example, \LaTeX~requires repeated rebuilding until it
reaches a fixed point, which can be directly expressed in build systems, such as
\Pluto~\cite{erdweg2015pluto}. Another example is \Excel, where a cell can
depend on itself, for example: \cmd{A1 = A1 + 1}. In such cases \Excel will
normally not execute anything, but if the ``Iterative Calculations'' feature is
enabled \Excel will execute the formula for a specified maximum number $N$ of
times per calculation (where $N$ is a setting that defaults to 100).

For examples like \LaTeX~we consider the proper encoding to not be circular
tasks, but a series of iterative steps, as described by
Mitchell~\shortcite{shake-fixed-point}. It is important that the number of
executions is bounded, otherwise the build system may not terminate (a
legitimate concern with \LaTeX, which can be put into a situation where it is
bistable or diverging over multiple executions). The examples in \Excel tend to
encode either mutable state, or recurrence relations. The former is only
required because \Excel inherently lacks the ability to write mutable state, and
the latter is probably better solved using explicit recurrence formulae.

We choose not to deal with cyclic dependencies -- a choice that most build
systems also follow. There are computation frameworks that support dependency
cycles under the assumption that tasks are \emph{monotonic} in a certain
sense~\cite{pottier2009lazy,radul2009propagation}.

\vspace{-1mm}
\subsection{Key-dependent Value Types}\label{sec-polymorphism}

Key-dependent value types allow a build system to work with multiple different
types of values, where the type of any particular value is determined by the
key. As an example of why this might be useful, consider a build system where
keys can be files (whose contents are strings) or system executables
(represented by their version number) -- using a single type for both values
reduces type safety. \Shake permits such key-dependent value types, e.g. see
the \emph{oracle} rule in Mitchell~\shortcite{mitchell2012shake}, and users
have remarked that this additional type safety provides a much easier expression
of concepts~\cite{hadrian}.

We can encode key-dependent value types using generalised algebraic data types,
or GADTs~\cite{spj2006gadts}. The idea is to replace
the callback \hs{fetch}~\hs{::}~\hs{k}~\hs{->}~\hs{f}~\hs{v} by its
more polymorphic equivalent \hs{fetch}~\hs{::}~\hs{k}~\hs{v}~\hs{->}~\hs{f}~\hs{v},
where~\hs{k}~\hs{v} is a GADT representing keys tagged by the type~\hs{v} of
corresponding values. The variable \hs{k} has changed from kind \hs{*} (a type),
to \hs{* -> *} (a type function), permitting the key to constrain the type of the value.
The idea is best explained by way of an example:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
data Version = Version { major :: Int, minor :: Int }
    deriving (Eq, Ord)

data Key a where
    File    :: FilePath -> Key String
    Program :: String -> Key Version
\end{minted}
\vspace{1mm}

\noindent
Here we extend the usual mapping from file paths to file contents with an
additional key type \hs{Program} which maps the name of a program to its installed \hs{Version}. The task
abstraction needs to be adjusted to cope with such keys (the suffix \hs{T}
stands for ``typed''):

\vspace{1mm}
\begin{minted}[xleftmargin=5pt]{haskell}
type Fetch k f = @\std{forall}@ v. k v -> f v

newtype TaskT c k v = TaskT (@\std{forall}@ f. c f => Fetch k f -> f v)

type TasksT c k = @\std{forall}@ v. k v -> Maybe (TaskT c k v)
\end{minted}
\vspace{1mm}

\noindent
The changes compared to the definition in~\S\ref{sec-task} are minimal: (i) the
\hs{TaskT} now uses a typed \hs{Fetch} callback (we define a separate type
synonym only for readability), and (ii) the type of \hs{TasksT} is now
polymorphic in \hs{v} instead of being parameterised by a concrete \hs{v}. The
example below demonstrates how \hs{fetch} can be used to retrieve dependencies
of different types: the rule \cmd{release.txt} concatenates the contents of two
\hs{File}s, while the rule \cmd{main.o} uses the numeric \hs{Program "gcc"} to
determine how the \hs{source} file should be compiled.

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
example :: TasksT Monad Key
example (File "release.txt") = Just $ TaskT $ \fetch -> do
    readme  <- fetch (File "README")
    license <- fetch (File "LICENSE")
    return (readme ++ license)
example (File "main.o") = Just $ TaskT $ \fetch -> do
    let source = "main.c"
    version <- fetch (Program "gcc")
    if version >= Version 8 0 then compileNew source
                              else compileOld source
example _ = Nothing
\end{minted}
\vspace{1mm}

Note that like all key-dependent tasks, this example could be expressed
without key-dependence, at the cost of some type safety. As we'll see in
\S\ref{sec-multiple-outputs} and \S\ref{sec-tracking-aspects}, using key-dependent
value types can make it easier to work with more complicated tasks.

\subsection{Multiple-Output Build Tasks}\label{sec-multiple-outputs}

Some build tasks produce multiple output keys -- for example \cmd{ghc A.hs} produces both
\cmd{A.hi} and \cmd{A.o}. This pattern can be encoded by having a key \cmd{A.hi+A.o} which produces
both results, then each of \cmd{A.hi} and \cmd{A.o} can extract the relevant result
from \cmd{A.hi+A.o}. We can express this pattern more clearly with the extra type safety
from~\S\ref{sec-polymorphism}:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
data Key a where
    File  :: FilePath   -> Key String
    Files :: [FilePath] -> Key [String]

task :: TasksT Monad Key
task (File "A.hi") = Just $ TaskT $ \fetch -> do
    [hi,o] <- fetch (Files ["A.hi","A.o"])
    return hi
task (File "A.o") = Just $ TaskT $ \fetch -> do
    [hi,o] <- fetch (Files ["A.hi","A.o"])
    return o
task (Files ["A.hi","A.o"]) = Just $ TaskT $ \fetch ->
    compileWithGHC "A.hs"
\end{minted}
\vspace{1mm}

One awkward aspect is that both \cmd{A.hi} and \cmd{A.o} must ask for
exactly the same \hs{Files} key. If one \hs{File} key swapped the order of the
list to \hs{Files} then it would likely run GHC twice. To help construct a
well-formed multiple-output build task it is convenient to partition the set of
keys into \emph{elementary subsets}. We can encode such a partition with a
function from any key to all the members of its subset (in a consistent order).

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
type Partition k = k -> [k]
\end{minted}
\vspace{1mm}

\noindent
In our example the partition function on either of the output names \hs{"A.hi"}
or \hs{"A.o"} would return \hs{["A.hi","A.o"]}.

With a suitable \hs{Partition} it is possible to create a mapping that resolves
\hs{File} keys automatically into the correct \hs{Files} key.

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
task :: Partition FilePath -> TasksT Monad Key
task partition (File k) = Just $ TaskT $ \fetch -> do
    let ks = partition k
    let Just i = elemIndex k ks
    vs <- fetch (Files ks)
    return (vs !! i)
task (Files ["A.hi","A.o"]) = Just $ TaskT $ \fetch ->
    compileWithGHC "A.hs"
... -- more tasks for elementary subsets as required
\end{minted}
\vspace{1mm}

In the above function we compute the elementary subset \hs{ks} of the given key
\hs{k}, find the index of \hs{k} within the subset (using \hs{elemIndex}), run
the \hs{fetch} to build every key in the subset, then extract out the value
corresponding to \hs{k} (using the indexing operation \hs{!!}).

% \todo{AM/NM}{We are talking about sets above, but are using lists, which is a
% bit inconsistent. We could switch from \hs{[k]} to \hs{Set k} fairly easily.
% Shall we? Neil says: Haskell code dealing with sets is quite a lot more complex.
% People are used to being quite fuzzy around the difference - I'd leave it as is.
% Andrey says: Also, using \hs{Set v} doesn't seem possible, because we won't be
% able to relate \hs{v}s with corresponding \hs{k}s}.

\subsection{Self-tracking}\label{sec-tracking-aspects}

Some build systems, for example \Excel and \Ninja, are capable of recomputing a
task if either its dependencies change, \emph{or} the task itself changes. For
example:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{text}
A1 = 20      B1 = A1 + A2
A2 = 10
\end{minted}
\vspace{1mm}

\noindent
In \Excel the user can alter the value produced by \cmd{B1} by either editing
the inputs of \cmd{A1} or \cmd{A2}, \emph{or} editing the formula in
\cmd{B1}~--~e.g. to \cmd{A1 * A2}. This pattern can be captured by describing
the rule producing \cmd{B1} as also depending on the value \cmd{B1-formula}.
The implementation can be given very directly in a
\hs{Tasks}~\hs{Monad}~--~concretely, first look up the formula, then interpret
it:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
sprsh6 "B1" = Just $ Task $ \fetch -> do
    formula <- fetch "B1-formula"
    evalFormula fetch formula
\end{minted}
\vspace{1mm}

\noindent
The build systems that have precise self-tracking are all ones which
use a \emph{non-embedded domain specific language} to describe build
tasks; that is, a task is described by a data structure (or syntax
tree), and tasks can be compared for equality by comparing those data
structures. Build systems that use a full programming language, e.g. \Shake,
are faced with the challenge of implementing equality on arbitrary
tasks -- and a task is just a function.  For such build systems, the
only safe approach is to assume (pessimistically) that any
change to the build system potentially changes any build task ~--~ the
classic example being build tasks depending on the makefile itself.

Below we show how to implement self-tracking in a build system that allows users
to describe build tasks by \emph{scripts} written in a non-embedded domain
specific language. We will denote the type of scripts by \hs{s}, and will assume
that scripts are indexed by keys \hs{k} just like all other values \hs{v}. More
specifically, we use the following GADT to tag keys \hs{k} with corresponding
result types: \hs{s} for scripts, and \hs{v} for other values.

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
data Key k v s a where
    Script :: k -> Key k v s s -- Keys for build scripts
    Value  :: k -> Key k v s v -- Keys for all other values
\end{minted}
\vspace{1mm}

\noindent
The function \hs{selfTracking} defined below is a generalisation of the approach
explained in the above \Excel example \hs{sprsh6}. The function takes a parser
for scripts, of type \hs{s}~\hs{->}~\hs{Task}~\hs{Monad}~\hs{k}~\hs{v}, and a
description of \emph{how to build all scripts}, of type
\hs{Tasks}~\hs{Monad}~\hs{k}~\hs{s}. For \hs{sprsh6}, the latter would simply
fetch \cmd{B1-formula} when given the key \cmd{B1} and return \hs{Nothing}
otherwise, but the presented approach can cope with much more sophisticated
scenarios where scripts themselves are derived from ``script sources'', e.g. all
C compilation scripts can be obtained from a single pattern rule, such as
\cmd{gcc -c [source] -o [object]}. The resulting typed task description
\hs{TasksT}~\hs{Monad}~\hs{(Key}~\hs{k}~\hs{v}~\hs{s)} tracks both values and
scripts that compute them, and is therefore self-tracking.

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
selfTracking :: (s -> Task Monad k v) -> Tasks Monad k s
                                      -> TasksT Monad (Key k v s)
selfTracking parse tasks key = case key of
    Script k -> getScript <$> tasks k
    Value  k -> runScript <$> tasks k
  where
    -- Build the script and return it
    getScript :: Task Monad k s -> TaskT Monad (Key k v s) s
    getScript task = TaskT $ \fetch -> run task (fetch . Script)
    -- Build the script, parse it, and then run the obtained task
    runScript :: Task Monad k s -> TaskT Monad (Key k v s) v
    runScript task = TaskT $ \fetch -> do
        script <- run task (fetch . Script)
        run (parse script) (fetch . Value)
\end{minted}
\vspace{1mm}

\noindent
It is possible to implement \hs{selfTracking} without relying on GADTs and typed
tasks presented in~\S\ref{sec-polymorphism}, at the cost of using partial
functions.

\subsection{File watching vs polling}\label{sec-file-watching}

Most build systems use files for their inputs. When running a build, to check if
a file is up-to-date, one option is to take the modification time or compute the
content hash of each input file. However, for the largest projects, the mere act
of checking the modification time for all the input files can become
prohibitive. To overcome that problem, build systems targeted at large scale,
e.g. \Bazel and \Buck, rely on file watching API's to detect when a file has
changed. Using that information the build can avoid performing many file
accesses.

File-watching build systems face new engineering challenges. Firstly, a running
task may become \emph{obsolete} due to a concurrent modification of its direct
or transitive dependencies. To minimise the amount of unnecessary work, a
scheduler can \emph{cancel} the obsolete task and restart it once its
dependencies are up-to-date. Note that topological
schedulers~(\S\ref{sec-topological}) cannot be easily adapted to support such
situations, since a new topological sort may be required when file dependencies
change. A further complication is that spurious dependency cycles may form
during a series of file modifications, and a file-watching build system
integrated with an IDE should be able to cope with such spurious cycles
gracefully instead of terminating with an error.