### 2019-10-20 papers/jfp/9-related.tex

\section{Related Work}\label{sec-related}

While there is research on individual build systems, there has been little
research to date comparing different build systems. In~\S\ref{sec-background} we
covered several important build systems~--~in this section we relate a few
other build systems to our abstractions, and discuss other work where similar
abstractions~arise.

\subsection{Other Build Systems}\label{sec-related-build}

Most build systems, when viewed at the level we talk, can be captured with minor
variations on the code presented in \S\ref{sec-implementations}. Below we list
some notable examples:

\begin{itemize}
\item \Dune~\cite{dune} is a build system designed for OCaml/Reason projects.
Its original implementation used \emph{arrows}~\cite{hughes2000generalising}
rather than monads to model dynamic dependencies, which simplified static
dependency approximation. \Dune was later redesigned to use a flavour of
selective functors~\cite{mokhov_selective_2019}, making it a closer fit to our
abstractions.

\item \Ninja~\cite{ninja} combines the \hs{topological} scheduler of \Make with
the verifying traces of \Shake~--~our associated implementation provides such a
combination. \Ninja~is also capable of modelling build rules that produce
multiple results, a limited form of multiple value types \S\ref{sec-polymorphism}.

\item \Nix~\cite{dolstra2004nix} has coarse-grained dependencies, with precise
hashing of dependencies and downloading of precomputed build products. We
provided a model of \Nix in \S\ref{sec-implementation-cloud}, although it is
worth noting that \Nix is not primarily intended as a build system, and the
coarse grained nature (packages, not individual files) makes it targeted to a
different purpose.

\item \Pluto~\cite{erdweg2015pluto} is based on a similar model to \Shake, but
additionally allows cyclic build rules combined with a user-specific resolution
strategy. Often such a strategy can be unfolded into the user rules without loss
of precision, but a fully general resolution handler extends the \hs{Task}
abstraction with new features.

\item \Redo~\cite{redo-idea,grosskurth2007redo,redo} almost exactly
matches \Shake at the level of detail given here, differing only on aspects like
rules producing multiple files~\S\ref{sec-polymorphism}. While \Redo predates
\Shake, they were developed independently; we use \Shake as a prototypical
example of a monadic build system because its implementation presents a closer
mapping to our \hs{Task} abstraction.

\item \Tup~\cite{tup} functions much like \Make, but with a refined dirty-bit
implementation that watches the file system for changes and can thus avoid
rechecking the entire graph. \Tup also automatically deletes stale results.
\end{itemize}

The one build system we are aware of that cannot be modelled in our framework is
\Fabricate by~Hoyt~\etal~\shortcite{fabricate}. In \Fabricate a build system is
a script that is run in-order, in the spirit of:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{bash}
gcc -c util.c
gcc -c main.c
gcc util.o main.o -o main.exe
\end{minted}
\vspace{1mm}

\noindent
To achieve minimality, each separate command is traced at the OS-level, allowing
\Fabricate to record a trace entry stating that \cmd{gcc -c util.c} reads from
\cmd{util.c}. In future runs \Fabricate runs the script from start to finish,
skipping any commands where no inputs have changed. The main difference from our
\hs{Tasks} abstraction is that instead of supplying a mapping from keys to
tasks, a \Fabricate script supplies a list of build statements, in a
user-scheduled order, without declaring what each statement reads or write.

Taking our abstraction, it is possible to encode \Fabricate assuming that
commands like \cmd{gcc -c util.c} are keys, there is a linear dependency between
each successive key, and that the OS-level tracing can be lifted back as a
monadic \hs{Task} function\footnote{\Shake provides support for
\Fabricate{}-like build systems~--~see \cmd{Development.Shake.Forward}.}.
However, in our pure model the mapping is not perfect as \cmd{gcc} writes to
arbitrary files whose locations are not known in advance. One way of capturing
arbitrary writes in our model is to switch from one callback \hs{fetch} to
\emph{two callbacks}, say \hs{read} and \hs{write}, allowing us to track both
reads and writes separately.

\subsection{Self-adjusting Computation}

While not typically considered build systems, self-adjusting computation is a
well studied area, and in particular the contrast between different formulations
has been thoroughly investigated, e.g.
see~Acar~\etal~\shortcite{acar2007selfadjusting}. Self-adjusting computations
can automatically adjust to an external change to their inputs. A classic
example is a self-adjusting sorting algorithm, which can efficiently (in
$O(\log{n})$ time where $n$ is the length of the input) recalculate the result
given an incremental change of the input. While very close to build systems in
spirit, self-adjusting computations are mostly used for in-memory computation
and rely on the ability to dynamically allocate new keys in the store for
sharing intermediate computations~--~an intriguing feature rarely seen in build
systems (\Shake's oracles~\S\ref{sec-polymorphism} can be used to model this
feature to a limited degree). Another important optimisation that self-adjusting
computation engines often support is the incremental processing of
\emph{deltas}, where instead of marking a value as ``changed to 8'', one can
mark it as ``changed by $+1$'', assuming it was equal to 7 before. When a delta
is small, it can often be propagated to the output more efficiently than by
recomputing the output value from scratch.

A lot of research has been dedicated to finding efficient data structures and
algorithms for self-adjusting computations, with a few open-source
implementations, e.g. \Incremental by Jane Street~\shortcite{incremental}. We
plan to investigate how these insights can be utilised by build systems as
future work.

\subsection{Memoization}\label{sec-related-memo}

\emph{Memoization} is a classic optimisation technique for storing values of a
function instead of recomputing them each time the function is called. Minimal
build systems (\S\ref{sec-background-make}) certainly perform
memoization: they \emph{store values instead of recomputing them each time}.
Memoization can therefore be reduced to a minimal build system (as we
demonstrate below), but not vice versa, since minimal build systems solve a more
complex optimisation problem.

As a simple example of using a build system for memoization, we solve a textbook
dynamic programming problem~--~Levenshtein's \emph{edit
distance}~\cite{levenshtein1966binary}: given two input strings $a$ and
$b$, find the shortest series of edit operations that transforms $a$
to $b$. The edit operations are typically \emph{inserting}, \emph{deleting} or
\emph{replacing} a symbol. The dynamic programming solution of this problem is
so widely known, e.g., see~Cormen~\etal~\shortcite{cormen2001introduction}, that
we provide its encoding in our \hs{Tasks} abstraction without further
explanation. We address elements of strings $a_i$ and $b_i$ by keys \hs{A}~$i$
and \hs{B}~$i$, respectively, while the cost of a subproblem $c_{ij}$ is
identified by \hs{C}~$i$~$j$.

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
data Key = A Int | B Int | C Int Int deriving Eq
\end{minted}
\begin{minted}[xleftmargin=10pt]{haskell}
editDistance :: Tasks Monad Key Int
editDistance (C i 0) = Just $ Task $ const $ pure i
editDistance (C 0 j) = Just $ Task $ const $ pure j
editDistance (C i j) = Just $ Task $ \fetch -> do
    ai <- fetch (A i)
    bj <- fetch (B j)
    if ai == bj
        then fetch (C (i - 1) (j - 1))
        else do insert  <- fetch (C  i      (j - 1))
                delete  <- fetch (C (i - 1)  j     )
                replace <- fetch (C (i - 1) (j - 1))
                return (1 + minimum [insert, delete, replace])
editDistance _ = Nothing
\end{minted}
\vspace{1mm}

\noindent
When asked to build \hs{C}~$n$~$m$, a minimal build system will calculate the
result using memoization. Furthermore, when an input $a_i$ is changed, only
necessary, incremental recomputation will be performed~--~an optimisation that
cannot be achieved just with memoization.

Self-adjusting computation, memoization and build systems are inherently related
topics, which poses the question of whether there is an underlying common
abstraction waiting to be discovered.